# Preprocess *d'*

## Helper function: Process item memory
```{r}
fun.summarize_item_block = function(df, group_vars) {
  df |> 
    filter(
      !trial_burnin,
      !resp_item_mem_acc %in% c("blink", "NO_RESPONSE")
    ) |> 
    group_by(across(all_of(group_vars))) |> 
    summarise(
      item_hit       = sum(resp_item_mem_acc == "Item Hit"),
      item_miss      = sum(resp_item_mem_acc == "Item Miss"),
      item_fa        = sum(resp_item_mem_acc == "Item FA"),
      item_cr        = sum(resp_item_mem_acc == "Item CR"),
      n_item_target  = sum(goal %in% c("BGSM", "PLUP")),
      n_item_foil    = sum(goal == "NOV"),
      n_item_total   = item_hit + item_miss + item_fa + item_cr,
      .groups        = "drop"
    ) |> 
    mutate(
      item_hr  = item_hit / n_item_target,
      item_far = item_fa  / n_item_foil,
      item_dp  = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil),
    )
}
```

## Helper function: Process source memory
```{r}
fun.summarize_src_block = function(df, group_vars) {
  df |> 
    filter(
      !trial_burnin,
      !resp_src_mem_acc %in% c("blink", "NO_RESPONSE")
    ) |> 
    group_by(across(all_of(group_vars))) |> 
    summarise(
      src_hit       = sum(resp_src_mem_acc %in% c("Size Hit", "Pleasant Hit")),
      src_miss      = sum(resp_src_mem_acc %in% c("Size Miss", "Pleasant Miss")),
      src_fa        = sum(resp_src_mem_acc %in% c("Size FA", "Pleasant FA")),
      src_cr        = sum(resp_src_mem_acc == "Source CR"),
      n_src_target  = sum(goal %in% c("BGSM", "PLUP")),
      n_src_foil    = sum(goal == "NOV"),
      n_src_total   = src_hit + src_miss + src_fa + src_cr,
      .groups       = "drop"
    ) |> 
    mutate(
      src_hr  = src_hit / n_src_target,
      src_far = src_fa  / n_src_foil,
      src_dp  = fun.dprime(src_hit, src_fa, n_src_target, n_src_foil),
    )
}
```

## Make Dataframes

### Load in trial-level beh data
```{r}
df.trials = read_tsv(
  here("data", paste0(path.bids_base, "trials_beh.tsv")),
  show_col_types = FALSE
)

df.trials |> 
  head(10) |> 
  kable(digits = 2)
```

### Item memory
```{r}
list.item_memory = list()
```

#### ...by subject
```{r}
df.item_dprime_by_subj = fun.summarize_item_block(
  df = df.trials,
  group_vars = "participant_id"
)

list.item_memory$by_subj = df.item_dprime_by_subj

df.item_dprime_by_subj |> 
  head(10) |> 
  kable(digits = 2)
```

#### ...by subject $\times$ realtime pupil status
```{r}
df.item_dprime_by_subj_x_realtime_pupil = fun.summarize_item_block(
  df = df.trials,
  group_vars = c(
    "participant_id", 
    "realtime_window_status"
  )
)

list.item_memory$by_subj_x_realtime_pupil = df.item_dprime_by_subj_x_realtime_pupil

df.item_dprime_by_subj_x_realtime_pupil |> 
  head(10) |> 
  kable(digits = 2)
```

#### ...by subject $\times$ realtime trigger condition
```{r}
df.item_dprime_by_subj_x_realtime_trigger = fun.summarize_item_block(
  df = df.trials,
  group_vars = c(
    "participant_id",
    "trigger_status"
  )
)

list.item_memory$by_subj_x_realtime_trigger = df.item_dprime_by_subj_x_realtime_trigger

df.item_dprime_by_subj_x_realtime_trigger |> 
  head(10) |> 
  kable(digits = 2)
```

#### ...by subject $\times$ realtime pupil status $\times$ trigger condition
```{r}
df.item_dprime_by_subj_x_realtime_pupil_x_trigger = fun.summarize_item_block(
  df.trials,
  group_vars = c(
    "participant_id",
    "realtime_window_status",
    "trigger_status"
  )
)

list.item_memory$by_subj_x_realtime_pupil_x_trigger = df.item_dprime_by_subj_x_realtime_pupil_x_trigger

df.item_dprime_by_subj_x_realtime_pupil_x_trigger |> 
  head(10) |> 
  kable(digits = 2)
```

#### ...by stimulus
```{r}
df.item_dprime_by_stim = fun.summarize_item_block(
  df = df.trials,
  group_vars = "stim"
)

list.item_memory$by_stim = df.item_dprime_by_stim

df.item_dprime_by_stim |> 
  head(10) |> 
  kable(digits = 2)
```

#### ...by goal state
```{r}
df.item_dprime_by_goal = fun.summarize_item_block(
  df = df.trials,
  group_vars = "goal"
)

list.item_memory$by_goal = df.item_dprime_by_goal

df.item_dprime_by_goal |> 
  head(10) |> 
  kable(digits = 2)
```

### Source memory
```{r}
list.src_memory = list()
```

#### ...by subject
```{r}
df.src_dprime_by_subj = fun.summarize_src_block(
  df = df.trials,
  group_vars = "participant_id"
)

list.src_memory$by_subj = df.src_dprime_by_subj

df.src_dprime_by_subj |> 
  head(10) |> 
  kable(digits = 2)
```

#### ...by subject $\times$ realtime pupil status
```{r}
df.src_dprime_by_subj_x_realtime_pupil = fun.summarize_src_block(
  df = df.trials,
  group_vars = c(
    "participant_id", 
    "realtime_window_status"
  )
)

list.src_memory$by_subj_x_realtime_pupil = df.src_dprime_by_subj_x_realtime_pupil

df.src_dprime_by_subj_x_realtime_pupil |> 
  head(10) |> 
  kable(digits = 2)
```

#### ...by subject $\times$ realtime trigger condition
```{r}
df.src_dprime_by_subj_x_realtime_trigger = fun.summarize_src_block(
  df = df.trials,
  group_vars = c(
    "participant_id",
    "trigger_status"
  )
)

list.src_memory$by_subj_x_realtime_trigger = df.src_dprime_by_subj_x_realtime_trigger

df.src_dprime_by_subj_x_realtime_trigger |> 
  head(10) |> 
  kable(digits = 2)
```

#### ...by subject $\times$ realtime pupil status $\times$ trigger condition
```{r}
df.src_dprime_by_subj_x_realtime_pupil_x_trigger = fun.summarize_src_block(
  df.trials,
  group_vars = c(
    "participant_id",
    "realtime_window_status",
    "trigger_status"
  )
)

list.src_memory$by_subj_x_realtime_pupil_x_trigger = df.src_dprime_by_subj_x_realtime_pupil_x_trigger

df.src_dprime_by_subj_x_realtime_pupil_x_trigger |> 
  head(10) |> 
  kable(digits = 2)
```

#### ...by stimulus
```{r}
df.src_dprime_by_stim = fun.summarize_src_block(
  df = df.trials,
  group_vars = "stim"
)

list.src_memory$by_stim = df.src_dprime_by_stim

df.src_dprime_by_stim |> 
  head(10) |> 
  kable(digits = 2)
```

#### ...by goal state
```{r}
df.src_dprime_by_goal = fun.summarize_src_block(
  df = df.trials,
  group_vars = "goal"
)

list.src_memory$by_goal = df.src_dprime_by_goal

df.src_dprime_by_goal |> 
  head(10) |> 
  kable(digits = 2)
```

## Outliers

### Identify problematic subjects

#### Item memory

Here, we identify any subjects with item memory *d'* less than or equal to 0 and/or any subjects who have very few or very many lapsing trials (i.e., more than 2 SDs from the mean proportion of lapsing trials relative to all subjects).

```{r}
lapse_item_counts = list.item_memory$by_subj_x_realtime_trigger |>
  mutate(
    trigger_status = factor(
      trigger_status, 
      levels = c(
        "no_lapse_no_trigger",
        "lapse_no_trigger",
        "lapse_trigger"
      )
    )
  ) |> 
  group_by(participant_id) |>
  summarise(
    lapse_sum = sum(n_item_total[trigger_status %in% c("lapse_no_trigger", "lapse_trigger")]),
    total_sum = sum(n_item_total),
    lapse_prop = lapse_sum / total_sum,
    any_dprime_zero_or_neg = any(item_dp <= 0)
  ) |> 
  mutate(
    z_score = scale(lapse_prop)[,1],
    is_extreme = abs(z_score) > 2 | any_dprime_zero_or_neg,  # more than 2 SD from mean
  )

lapse_item_counts |> 
  filter(is_extreme) |> 
  kable(digits = 3)
```

#### Source Memory

Again, we identify any subjects with source memory *d'* less than or equal to 0 and/or any subjects who have very few or very many lapsing trials (i.e., more than 2 SDs from the mean proportion of lapsing trials relative to all subjects).

-   note, the proportion of trials calculation is the same as above (for the item memory *d'* calculations, since the trial counts don't change based on their item vs. source memory)

```{r}
lapse_src_counts = list.src_memory$by_subj_x_realtime_trigger |>
  mutate(
    trigger_status = factor(
      trigger_status, 
      levels = c(
        "no_lapse_no_trigger",
        "lapse_no_trigger",
        "lapse_trigger"
      )
    )
  ) |>
  group_by(participant_id) |>
  summarise(
    lapse_sum = sum(n_src_total[trigger_status %in% c("lapse_no_trigger", "lapse_trigger")]),
    total_sum = sum(n_src_total),
    lapse_prop = lapse_sum / total_sum,
    any_dprime_zero_or_neg = any(src_dp <= 0)
  ) |> 
  mutate(
    z_score = scale(lapse_prop)[,1],
    is_extreme = abs(z_score) > 2 | any_dprime_zero_or_neg,  # more than 2 SD from mean
  )

lapse_src_counts |>
  filter(is_extreme) |> 
  kable(digits = 3)
```

### Failure to Comply with Task Instructions

-   according to experimenter logs, sub-27 misunderstood the instructions and intentionally withheld responses on all `lapse_trigger` trials, thus, they have no usable data here

```{r}
missing_conds = "sub-27"
```

-   other subjects to exclude based on experimenter logs

```{r}
bad_subjs = c("sub-78", "sub-80")
```

### Collapse Subject IDs

```{r}
ids_to_exclude = lapse_src_counts |> 
  filter(is_extreme) |> 
  pull(participant_id) |> 
  c(lapse_item_counts |> filter(is_extreme) |> pull(participant_id)) |>
  c(missing_conds, bad_subjs) |> 
  unique() |> 
  sort()

print(ids_to_exclude)

length(ids_to_exclude)
```

### Broken Down by Exclusion Criteria

```{r}
message("item memory d' (any less than or equal to 0):")
lapse_item_counts |>
  filter(is_extreme, any_dprime_zero_or_neg) |> 
  kable(digits = 3)

message("item memory d' (any more than 2 SDs from mean proportion of lapse trials:")
lapse_item_counts |>
  filter(is_extreme, !any_dprime_zero_or_neg) |> 
  kable(digits = 3)

message("source memory d' (any less than or equal to 0):")
lapse_src_counts |>
  filter(is_extreme, any_dprime_zero_or_neg) |> 
  kable(digits = 3)

message("source memory d' (any more than 2 SDs from mean proportion of lapse trials:")
lapse_src_counts |>
  filter(is_extreme, !any_dprime_zero_or_neg) |> 
  kable(digits = 3)
```

## Write Data

```{r}
fun.write_block(list.item_memory, "item_dprimes")
fun.write_block(list.src_memory, "source_dprimes")
fun.write_block(ids_to_exclude, "ids_to_exclude", skip_list = TRUE)
```
