# Spline detrending

## Parameters
```{r}
SPLINE_DF = 5
OUTLIER_THRESHOLD = 1.0
INCLUDE_BURNIN_IN_BASELINE = TRUE
CLASSIFY_BURNIN_TRIALS = FALSE
EPOCH_ORDER = c("fixprestim", "realtime1", "realtime2", "trigger", "notrigger", "fixposttrig", "stimresp")
```

## Load exclusion list
```{r}
vec.subjects_to_exclude <- ids_to_exclude
```

## Get list of parquet files
```{r}
parquet_files = list.files(
  here(path.pupil_dir, "ses-ret"),
  pattern = "*_desc-beh_pupil_preproc.parquet",
  full.names = TRUE
)

message(paste("Found", length(parquet_files), "parquet files"))
```

## Load trial-level behavioral data
```{r}
# Save stimulus-to-stim_index mapping before filtering (needed for chronological ordering)
df.trial_order <- df.trials |>
  select(participant_id, run, stim, stim_index) |>
  distinct()

df.trials <- df.trials |>
  filter(!participant_id %in% vec.subjects_to_exclude) |>
  filter(!resp_item_mem_acc %in% c("blink", "NO_RESPONSE"))

message(paste("Loaded", nrow(df.trials), "trials"))
```

## Spline detrending function
```{r}
detrend_with_spline = function(df, spline_df = SPLINE_DF, trial_order = df.trial_order) {
  df = as.data.frame(df)
  df = df[df$epoch %in% EPOCH_ORDER, ]
  df$epoch_rank = match(df$epoch, EPOCH_ORDER)

  # Get chronological stim_index from behavioral data
  df$stim_index <- NULL
  if (is.numeric(df$subject_id)) {
    df$subject_id <- paste0("sub-", str_pad(df$subject_id, width = 2, pad = "0"))
  }
  df <- df |>
    left_join(
      trial_order |>
        select(participant_id, run, stim, stim_index) |>
        rename(item = stim) |>
        distinct(),
      by = c("subject_id" = "participant_id", "run", "item")
    )

  # Arrange chronologically
  df <- df |>
    arrange(subject_id, run, stim_index, epoch_rank, timebin)

  # Compute within-trial sample position
  df <- df |>
    group_by(subject_id, run, stim_index) |>
    mutate(within_trial_sample = row_number() - 1) |>
    ungroup()

  # Median trial duration (in samples) per run — used to size gaps for missing trials
  trial_durations <- df |>
    count(subject_id, run, stim_index) |>
    group_by(subject_id, run) |>
    summarise(median_trial_samples = median(n), .groups = "drop")

  # Compute time_sec: stim_index sets the trial offset, gaps preserved for missing indices
  df <- df |>
    left_join(trial_durations, by = c("subject_id", "run")) |>
    mutate(time_sec = stim_index * median_trial_samples / 1000 + within_trial_sample / 1000)

  # Fit spline and detrend
  df |>
    group_by(subject_id, run) |>
    mutate(
      spline_trend = predict(
        lm(pupil ~ splines::ns(time_sec, df = spline_df), na.action = na.exclude)
      ),
      pupil_detrended = pupil - spline_trend
    ) |>
    ungroup() |>
    select(-within_trial_sample, -median_trial_samples) |>
    mutate(participant_id = paste0("sub-", str_pad(subject_id, width = 2, pad = "0"))) 
}
```

## Subject-level preproc function
```{r}
process_subject_spline = function(parquet_path, subjects_to_exclude, spline_df = SPLINE_DF, outlier_threshold = OUTLIER_THRESHOLD,
                                  include_burnin_in_baseline = INCLUDE_BURNIN_IN_BASELINE, classify_burnin_trials = CLASSIFY_BURNIN_TRIALS) {
  df.pupil = read_parquet(parquet_path)
  subj_id = unique(df.pupil$subject_id)[1]

  if (subj_id %in% subjects_to_exclude) {
    return(NULL)
  }

  # Exclude sub-71 run 3 (removed from df.trials due to data issue)
  df.pupil <- df.pupil |> filter(!(subject_id == subj_id & subj_id %in% c("sub-71", 71) & run == 3))

  # step 1: apply spline detrending
  df.pupil = detrend_with_spline(df.pupil, spline_df = spline_df)
  
  # step 2: extract per-trial baseline from fixprestim
  df.trial_baselines = df.pupil |> 
    filter(epoch == "fixprestim") |> 
    group_by(subject_id, run, trial, trial_burnin) |> 
    arrange(timebin, .by_group = TRUE) |> 
    slice_tail(n = 2000) |> 
    summarize(
      trial_baseline_detrended = mean(pupil_detrended, na.rm = TRUE),
      trial_baseline_raw = mean(pupil, na.rm = TRUE),
      .groups = "drop"
    )
  
  # step 3: get trial-level metadata
  df.trial_meta = df.pupil |>
    group_by(subject_id, run, trial, trial_burnin) |>
    summarize(
      stim_index = first(stim_index),
      trigger_lower_threshold = first(trigger_lower_threshold),
      trigger_upper_threshold = first(trigger_upper_threshold),
      trigger_baseline_mean = first(trigger_baseline_mean),
      trigger_baseline_std = first(trigger_baseline_std),
      trigger_status_original = first(trigger_status),
      trigger_actually_delivered = first(trigger_actually_delivered),
      trigger_assay_one_mean = first(trigger_assay_one_mean),
      trigger_assay_two_mean = first(trigger_assay_two_mean),
      .groups = "drop"
    )
  
  # step 4: compute realtime pupil value (detrended)
  df.realtime = df.pupil |> 
    filter(epoch %in% c("realtime1", "realtime2")) |> 
    group_by(subject_id, run, trial) |> 
    summarize(
      realtime_pupil_detrended = mean(pupil_detrended, na.rm = TRUE),
      realtime_pupil_raw = mean(pupil, na.rm = TRUE),
      .groups = "drop"
    )
  
  # step 5: merge all data
  df.trial_summary <- df.trial_meta |> 
    left_join(df.trial_baselines, by = c("subject_id", "run", "trial", "trial_burnin")) |> 
    left_join(df.realtime, by = c("subject_id", "run", "trial"))
  
  # step 6: compute global baseline stats
  df.for_baseline = df.trial_summary
  if (!include_burnin_in_baseline) {
    df.for_baseline = df.trial_summary |> 
      filter(!trial_burnin)
  }
  
  df.global_stats = df.for_baseline |> 
    group_by(subject_id, run) |> 
    summarize(
      global_baseline_mean = mean(trial_baseline_detrended, na.rm = TRUE),
      global_baseline_std = sd(trial_baseline_detrended, na.rm = TRUE),
      global_realtime_mean = mean(realtime_pupil_detrended, na.rm = TRUE),
      global_realtime_sd = sd(realtime_pupil_detrended, na.rm = TRUE),
      n_trials_in_baseline = n(),
      .groups = "drop"
    )
  
  # step 7: join global stats and compute thresholds
  df.trial_summary = df.trial_summary |> 
    left_join(df.global_stats, by = c("subject_id", "run")) |> 
    mutate(
      global_upper_threshold = global_realtime_mean + (outlier_threshold * global_realtime_sd),
      global_lower_threshold = global_realtime_mean - (outlier_threshold * global_realtime_sd)
    )
  
  # step 8: classify trials based on global thresholds
  if (classify_burnin_trials) {
    # classify all trials including burn-in
    df.trial_summary = df.trial_summary |> 
      mutate(
        is_lapse_global = case_when(
          realtime_pupil_detrended > global_upper_threshold |
            realtime_pupil_detrended < global_lower_threshold ~ TRUE,
          TRUE ~ FALSE
        ),
        trigger_status_new = case_when(
          is_lapse_global & trigger_actually_delivered ~ "lapse_trigger",
          is_lapse_global & !trigger_actually_delivered ~ "lapse_notrigger",
          !is_lapse_global & !trigger_actually_delivered ~ "nolapse_notrigger",
          !is_lapse_global & trigger_actually_delivered ~ "nolapse_trigger"
        )
      )
  } else {
    # only classify non-burnin trials
    df.trial_summary = df.trial_summary |> 
      mutate(
        is_lapse_global = case_when(
          trial_burnin ~ NA,
          realtime_pupil_detrended > global_upper_threshold |
            realtime_pupil_detrended < global_lower_threshold ~ TRUE,
          TRUE ~ FALSE
        ),
        trigger_status_new = case_when(
          is.na(is_lapse_global) ~ NA_character_,
          is_lapse_global & trigger_actually_delivered ~ "lapse_trigger",
          is_lapse_global & !trigger_actually_delivered ~ "lapse_notrigger",
          !is_lapse_global & !trigger_actually_delivered ~ "nolapse_notrigger",
          !is_lapse_global & trigger_actually_delivered ~ "nolapse_trigger"
        )
      )
  }
  
  rm(df.pupil)
  gc()
  
  return(df.trial_summary)
}
```

## test out on single subject
```{r fig.width=10, fig.height=15}
example_file = parquet_files[1]
df.example = read_parquet(example_file)
df.example_detrended = detrend_with_spline(df.example, spline_df = SPLINE_DF)

df.plot_spline <- df.example_detrended |> 
  filter(run == 1) |> 
  filter(row_number() %% 100 == 1)

detrend_p1 <- ggplot(df.plot_spline, aes(x = time_sec)) + 
  geom_line(aes(y = pupil), color = "gray50", alpha = 0.5) +
  geom_line(aes(y = spline_trend), color = "red", linewidth = 1.5) +
  labs(
    title = "original pupil + spline trend",
    x = "time (s)",
    y = "pupil size"
  )

detrend_p2 <- ggplot(df.plot_spline, aes(x = time_sec)) + 
  geom_line(aes(y = pupil_detrended), color = "steelblue", alpha = 0.7) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "detrended pupil (residuals)",
    x = "time (s)",
    y = "detrended pupil"
  )

detrend_p1 / detrend_p2
```

## Process all subjects
```{r}
results_list <- list()

message(paste("Processing", length(parquet_files), "subjects..."))

for (i in seq_along(parquet_files)) {
  f <- parquet_files[i]
  subj <- str_extract(basename(f), "sub-[0-9]+")
  
  if (i %% 5 == 0 || i == 1) {
    message(paste0("  Processing ", i, "/", length(parquet_files), ": ", subj))
  }
  
  result <- tryCatch(
    process_subject_spline(f, vec.subjects_to_exclude, 
                           spline_df = SPLINE_DF, 
                           outlier_threshold = OUTLIER_THRESHOLD,
                           include_burnin_in_baseline = INCLUDE_BURNIN_IN_BASELINE,
                           classify_burnin_trials = CLASSIFY_BURNIN_TRIALS),
    error = function(e) {
      message(paste("  ERROR processing", subj, ":", e$message))
      return(NULL)
    }
  )
  
  if (!is.null(result)) {
    results_list[[length(results_list) + 1]] <- result
  }
  
  if (i %% 10 == 0) {
    gc()
  }
}

message("Done processing all subjects.")
```

```{r}
df.reclassified <- bind_rows(results_list)
# rm(results_list); gc(verbose = FALSE)

message(glue("Subjects: {n_distinct(df.reclassified$subject_id)} | Trials: {nrow(df.reclassified)}"))

# Create comparison df with standardized 4-category original status
df.comparison <- df.reclassified |>
  mutate(
    trigger_status_original_4cat = case_when(
      trigger_status_original == "lapse_trigger" ~ "lapse_trigger",
      trigger_status_original == "lapse_no_trigger" ~ "lapse_notrigger",
      trigger_status_original == "no_lapse_no_trigger" ~ "nolapse_notrigger",
      trigger_status_original == "burnin" ~ "burnin",
      TRUE ~ trigger_status_original
    ),
    classification_changed = case_when(
      trial_burnin ~ NA,
      is.na(trigger_status_new) ~ NA,
      TRUE ~ trigger_status_original_4cat != trigger_status_new
    )
  )

# Merge reclassification into behavioral trial data
df.trials_reclassified <- df.trials |>
  left_join(
    df.comparison |>
      select(subject_id, run, trial, trigger_status_new,
             trigger_status_original_4cat, classification_changed) |>
      rename(participant_id = subject_id),
    by = c("participant_id", "run", "trial")
  )

message(glue("Merged: {nrow(df.trials_reclassified)} trials | With classification: {sum(!is.na(df.trials_reclassified$trigger_status_new))}"))
```

## Trial counts
```{r}
# Reclassification flow table: Original → Spline
message("=== Trial Flow: Original \u2192 Spline ===\n")

flow_table <- df.trials_reclassified |>
  filter(!trial_burnin, !is.na(trigger_status_new)) |>
  count(trigger_status_original_4cat, trigger_status_new) |>
  pivot_wider(names_from = trigger_status_new, values_from = n, values_fill = 0)

flow_table
```

```{r}
# Reclassification flow table split by constriction / dilation
message("=== Trial Flow by Constriction / Dilation ===\n")

flow_by_direction <- df.trials_reclassified |>
  filter(!trial_burnin, !is.na(trigger_status_new),
         realtime_window_status %in% c("constriction", "dilation")) |>
  count(trigger_status_original_4cat, trigger_status_new, realtime_window_status) |>
  pivot_wider(
    names_from = realtime_window_status,
    values_from = n,
    values_fill = 0
  ) |>
  mutate(cell_label = paste0(constriction, " / ", dilation)) |>
  select(trigger_status_original_4cat, trigger_status_new, cell_label) |>
  pivot_wider(
    names_from = trigger_status_new,
    values_from = cell_label,
    values_fill = "0 / 0"
  )

flow_by_direction
```

```{r}
df.trials_reclassified |>
  filter(!trial_burnin, !is.na(trigger_status_original_4cat)) |>
  count(trigger_status_original_4cat) |>
  mutate(pct = round(100 * n / sum(n), 1))

# Trial distribution by spline classification

df.trials_reclassified |>
  filter(!trial_burnin, !is.na(trigger_status_new)) |>
  count(trigger_status_new) |>
  mutate(pct = round(100 * n / sum(n), 1))
```

```{r}
# Per-subject trial counts per condition (summary stats)
message("=== Per-Subject Trial Counts ===\n")

df.trials_reclassified |>
  filter(!is.na(trigger_status_new), !trial_burnin) |>
  count(participant_id, trigger_status_new) |>
  group_by(trigger_status_new) |>
  summarise(
    n_subjects = n(),
    min_trials = min(n),
    median_trials = median(n),
    mean_trials = round(mean(n), 1),
    max_trials = max(n),
    .groups = "drop"
  )
```

## Spline Trend Spaghetti Plot
```{r}
message("Extracting epoch-stitched spline trends for spaghetti plot...")

n_files <- length(parquet_files)
spline_trends <- map_dfr(seq_along(parquet_files), function(i) {
  f <- parquet_files[i]
  subj <- str_extract(basename(f), "sub-[0-9]+")
  if (i %% 10 == 0 || i == 1 || i == n_files) message(glue("  [{i}/{n_files}] {subj}"))
  if (subj %in% vec.subjects_to_exclude) return(NULL)

  read_parquet(f) |>
    as.data.frame() |>
    filter(!(subject_id %in% c("sub-71", 71) & run == 3)) |>
    detrend_with_spline(spline_df = SPLINE_DF) |>
    select(subject_id, run, time_sec, spline_trend) |>
    group_by(subject_id, run) |>
    filter(row_number() %% 500 == 1) |>
    ungroup()
})

message(glue("Subjects: {n_distinct(spline_trends$subject_id)} | ",
             "Runs: {n_distinct(interaction(spline_trends$subject_id, spline_trends$run))}"))
```

```{r}
# Mean-center per subject-run
spline_trends <- spline_trends |>
  mutate(subj_run = interaction(subject_id, run)) |>
  group_by(subj_run) |>
  mutate(spline_trend = spline_trend - mean(spline_trend, na.rm = TRUE)) |>
  ungroup()

spline_range <- spline_trends |>
  group_by(subject_id, run, subj_run) |>
  summarise(trend_range = max(spline_trend) - min(spline_trend), .groups = "drop")

range_threshold <- median(spline_range$trend_range) + 2 * mad(spline_range$trend_range)
outlier_subj_runs <- filter(spline_range, trend_range > range_threshold)
outlier_ids <- unique(outlier_subj_runs$subject_id)

spline_trends <- spline_trends |>
  mutate(is_outlier = subj_run %in% outlier_subj_runs$subj_run)

message(glue("Range threshold: {round(range_threshold, 1)} | ",
             "Outlier subject-runs: {nrow(outlier_subj_runs)} from {length(outlier_ids)} subjects"))
```

```{r}
plot.spline_spaghetti <- ggplot(spline_trends, aes(x = time_sec, y = spline_trend, group = subj_run)) +
  geom_line(data = spline_trends,
            alpha = 0.3, linewidth = 1.2, color = "steelblue") +
  geom_line(data = spline_trends,
            aes(color = subject_id), alpha = 0.8, linewidth = 1.2) +
  labs(
    # title = "Mean-Centered Spline Trends Across All Subjects/Runs",
    x = "Time (seconds)",
    y = "Spline Trend",# (mean-centered)",
    color = "Outlier"
  ) +
  theme(legend.position = "none")

plot.spline_spaghetti
```

## Compute d' per reclassified condition

```{r}
spline_cond_levels = c("nolapse_notrigger", "lapse_notrigger", "lapse_trigger")
spline_cond_labels = c("NoLapse", "Lapse", "LapseTrigger")

# Item d' by subject x reclassified condition
df.spline_item_dp <- df.trials_reclassified |>
  filter(!trial_burnin, !is.na(trigger_status_new)) |>
  filter(trigger_status_new %in% spline_cond_levels) |>
  group_by(participant_id, trigger_status_new) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_item_target = sum(goal %in% c("BGSM", "PLUP")),
    n_item_foil = sum(goal == "NOV"),
    .groups = "drop"
  ) |>
  mutate(
    item_dp = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil),
    condition = factor(trigger_status_new, levels = spline_cond_levels, labels = spline_cond_labels)
  )

# Source d' by subject x reclassified condition
df.spline_source_dp <- df.trials_reclassified |>
  filter(!trial_burnin, !is.na(trigger_status_new)) |>
  filter(trigger_status_new %in% spline_cond_levels) |>
  filter(!resp_src_mem_acc %in% c("blink", "NO_RESPONSE")) |>
  group_by(participant_id, trigger_status_new) |>
  summarise(
    src_hit = sum(resp_src_mem_acc %in% c("Size Hit", "Pleasant Hit")),
    src_fa = sum(resp_src_mem_acc %in% c("Size FA", "Pleasant FA")),
    n_src_target = sum(goal %in% c("BGSM", "PLUP")),
    n_src_foil = sum(goal == "NOV"),
    .groups = "drop"
  ) |>
  mutate(
    src_dp = fun.dprime(src_hit, src_fa, n_src_target, n_src_foil),
    condition = factor(trigger_status_new, levels = spline_cond_levels, labels = spline_cond_labels)
  )

message(glue("Item d': {nrow(df.spline_item_dp)} rows from {n_distinct(df.spline_item_dp$participant_id)} subjects"))
message(glue("Source d': {nrow(df.spline_source_dp)} rows from {n_distinct(df.spline_source_dp$participant_id)} subjects"))
```

## Bayesian model: Item d' (spline reclassified)

```{r fig.height=20, warning=FALSE}
fit.brm_spline_item_dprime = brm(
  item_dp ~ 0 + condition,
  data    = df.spline_item_dp,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_spline_item_dprime"
)

fit.brm_spline_item_dprime |> summary()
fit.brm_spline_item_dprime |> plot(nvariables = 8)
fit.brm_spline_item_dprime |> pp_check(ndraws = 100)

fit.brm_spline_item_dprime |>
  as_draws_df() |>
  select(starts_with("b_"), sigma) |>
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") |>
  ggplot(mapping = aes(y = fct_rev(variable), x = value)) +
  stat_halfeye(fill = "lightblue") +
  geom_vline(xintercept = 0.0, linetype = "dashed") +
  theme(axis.title.y = element_blank())

# Lapse < NoLapse
hypothesis(fit.brm_spline_item_dprime, "conditionLapse < conditionNoLapse")

# LapseTrigger > Lapse
hypothesis(fit.brm_spline_item_dprime, "conditionLapseTrigger > conditionLapse")

# LapseTrigger < NoLapse
hypothesis(fit.brm_spline_item_dprime, "conditionLapseTrigger < conditionNoLapse")

fit.brm_spline_item_dprime |>
  emmeans(pairwise ~ condition)
```

```{r fig.width=6, fig.height=5, fig.dpi=300}
contrast_draws <- fit.brm_spline_item_dprime %>%
  spread_draws(b_conditionNoLapse, b_conditionLapse, b_conditionLapseTrigger) %>%
  mutate(
    `Lapse – No Lapse`    = b_conditionLapse - b_conditionNoLapse,
    `Trigger – Lapse`     = b_conditionLapseTrigger - b_conditionLapse,
    `No Lapse – Trigger`  = b_conditionNoLapse - b_conditionLapseTrigger
  ) %>%
  pivot_longer(cols = matches("\u2013"), names_to = "contrast", values_to = ".value") %>%
  mutate(contrast = factor(contrast, levels = c(
    "No Lapse \u2013 Trigger", "Trigger \u2013 Lapse", "Lapse \u2013 No Lapse"
  )))

fill_by_type <- c(
  "Lapse \u2013 No Lapse"   = "#D62729",
  "Trigger \u2013 Lapse"    = "#F67E00",
  "No Lapse \u2013 Trigger" = "#FCBF4A"
)

plot.spline_item_memory_contrasts = contrast_draws %>%
  ggplot(aes(x = .value, y = contrast, fill = contrast)) +
  stat_halfeye(point_interval = median_hdi, .width = c(0.66, 0.9),
               slab_alpha = 0.8, normalize = "panels", point_size = 4) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66, color = "black") +
  geom_text(
    data = contrast_draws %>% group_by(contrast) %>% summarise(m = median(.value)),
    aes(x = m, y = contrast, label = contrast),
    vjust = -0.8, hjust = 0.5, size = 6, inherit.aes = FALSE, family = "Inter"
  ) +
  scale_fill_manual(values = fill_by_type) +
  scale_y_discrete(labels = NULL) +
  labs(x = "Posterior contrast (d\u2032)", y = NULL, title = "Item memory (reclassified)") +
  theme(
    legend.position = "none", panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(), axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

plot.spline_item_memory_contrasts
```

## Bayesian model: Source d' (spline reclassified)

```{r fig.height=20, warning=FALSE}
fit.brm_spline_source_dprime = brm(
  src_dp ~ 0 + condition,
  data    = df.spline_source_dp,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_spline_source_dprime"
)

fit.brm_spline_source_dprime |> summary()
fit.brm_spline_source_dprime |> plot(nvariables = 8)
fit.brm_spline_source_dprime |> pp_check(ndraws = 100)

fit.brm_spline_source_dprime |>
  as_draws_df() |>
  select(starts_with("b_"), sigma) |>
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") |>
  ggplot(mapping = aes(y = fct_rev(variable), x = value)) +
  stat_halfeye(fill = "lightblue") +
  geom_vline(xintercept = 0.0, linetype = "dashed") +
  theme(axis.title.y = element_blank())

# Lapse < NoLapse
hypothesis(fit.brm_spline_source_dprime, "conditionLapse < conditionNoLapse")

# LapseTrigger > Lapse
hypothesis(fit.brm_spline_source_dprime, "conditionLapseTrigger > conditionLapse")

# LapseTrigger = NoLapse (two-sided)
hypothesis(fit.brm_spline_source_dprime, "conditionLapseTrigger < conditionNoLapse")

fit.brm_spline_source_dprime |>
  emmeans(pairwise ~ condition)
```

```{r fig.width=6, fig.height=5, fig.dpi=300}
contrast_draws <- fit.brm_spline_source_dprime %>%
  spread_draws(b_conditionNoLapse, b_conditionLapse, b_conditionLapseTrigger) %>%
  mutate(
    `Lapse – No Lapse`    = b_conditionLapse - b_conditionNoLapse,
    `Trigger – Lapse`     = b_conditionLapseTrigger - b_conditionLapse,
    `No Lapse – Trigger`  = b_conditionNoLapse - b_conditionLapseTrigger
  ) %>%
  pivot_longer(cols = matches("\u2013"), names_to = "contrast", values_to = ".value") %>%
  mutate(contrast = factor(contrast, levels = c(
    "No Lapse \u2013 Trigger", "Trigger \u2013 Lapse", "Lapse \u2013 No Lapse"
  )))

fill_by_type <- c(
  "Lapse \u2013 No Lapse"   = "#D62729",
  "Trigger \u2013 Lapse"    = "#F67E00",
  "No Lapse \u2013 Trigger" = "#FCBF4A"
)

plot.spline_source_memory_contrasts = contrast_draws %>%
  ggplot(aes(x = .value, y = contrast, fill = contrast)) +
  stat_halfeye(point_interval = median_hdi, .width = c(0.66, 0.9),
               slab_alpha = 0.8, normalize = "panels", point_size = 4) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66, color = "black") +
  geom_text(
    data = contrast_draws %>% group_by(contrast) %>% summarise(m = median(.value)),
    aes(x = m, y = contrast, label = contrast),
    vjust = -0.8, hjust = 0.5, size = 6, inherit.aes = FALSE, family = "Inter"
  ) +
  scale_fill_manual(values = fill_by_type) +
  scale_y_discrete(labels = NULL) +
  labs(x = "Posterior contrast (d\u2032)", y = NULL, title = "Source memory (reclassified)") +
  theme(
    legend.position = "none", panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(), axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

plot.spline_source_memory_contrasts
```

## "True lapses" trigger effect

```{r}
# Trials classified as lapse under BOTH original and spline methods
df.true_lapses <- df.trials_reclassified |>
  filter(!trial_burnin, !is.na(trigger_status_new)) |>
  filter(
    trigger_status_original_4cat %in% c("lapse_trigger", "lapse_notrigger"),
    trigger_status_new %in% c("lapse_trigger", "lapse_notrigger")
  )

true_lapse_cond_levels = c("lapse_notrigger", "lapse_trigger")
true_lapse_cond_labels = c("Lapse", "LapseTrigger")

# Item d' for true lapses
df.true_lapse_item_dp <- df.true_lapses |>
  group_by(participant_id, trigger_status_new) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_item_target = sum(goal %in% c("BGSM", "PLUP")),
    n_item_foil = sum(goal == "NOV"),
    n_trials = n(),
    .groups = "drop"
  ) |>
  mutate(
    item_dp = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil),
    condition = factor(trigger_status_new, levels = true_lapse_cond_levels, labels = true_lapse_cond_labels)
  )

message(glue("True lapses - Item d': {nrow(df.true_lapse_item_dp)} rows from {n_distinct(df.true_lapse_item_dp$participant_id)} subjects"))
df.true_lapse_item_dp |> count(condition)
```

```{r warning=FALSE}
fit.brm_true_lapse_item = brm(
  item_dp ~ 0 + condition,
  data    = df.true_lapse_item_dp,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_true_lapse_item_dprime"
)

fit.brm_true_lapse_item |> summary()

# LapseTrigger > Lapse (trigger rescue among true lapses)
hypothesis(fit.brm_true_lapse_item, "conditionLapseTrigger > conditionLapse")

fit.brm_true_lapse_item |> emmeans(pairwise ~ condition)
```

## Asymmetric trigger effect: Constriction vs. Dilation

```{r}
# Split true lapses by pupil direction (constriction = low arousal, dilation = high arousal)
df.true_lapses_by_direction <- df.true_lapses |>
  filter(realtime_window_status %in% c("constriction", "dilation"))

message(glue("True lapses with direction: {nrow(df.true_lapses_by_direction)} / {nrow(df.true_lapses)} total"))
df.true_lapses_by_direction |> count(trigger_status_new, realtime_window_status)
```

```{r}
# Item d' per subject × trigger condition × pupil direction
df.direction_item_dp <- df.true_lapses_by_direction |>
  group_by(participant_id, trigger_status_new, realtime_window_status) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_item_target = sum(goal %in% c("BGSM", "PLUP")),
    n_item_foil = sum(goal == "NOV"),
    n_trials = n(),
    .groups = "drop"
  ) |>
  mutate(
    item_dp = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil),
    trigger = factor(
      trigger_status_new,
      levels = c("lapse_notrigger", "lapse_trigger"),
      labels = c("NoTrigger", "Trigger")
    ),
    direction = factor(realtime_window_status, levels = c("constriction", "dilation"))
  )

message(glue("Direction × Trigger: {nrow(df.direction_item_dp)} rows from {n_distinct(df.direction_item_dp$participant_id)} subjects"))
df.direction_item_dp |> count(trigger, direction)
```

```{r warning=FALSE}
# 2×2 model: trigger × pupil direction
fit.brm_direction_item = brm(
  item_dp ~ 0 + trigger:direction,
  data    = df.direction_item_dp,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_direction_trigger_item_dprime"
)

fit.brm_direction_item |> summary()

# Trigger effect within constriction trials
hypothesis(fit.brm_direction_item,
  "triggerTrigger:directionconstriction > triggerNoTrigger:directionconstriction")

# Trigger effect within dilation trials
hypothesis(fit.brm_direction_item,
  "triggerTrigger:directiondilation > triggerNoTrigger:directiondilation")

# Asymmetry: trigger effect on constriction > trigger effect on dilation
hypothesis(fit.brm_direction_item,
  "(triggerTrigger:directionconstriction - triggerNoTrigger:directionconstriction) > (triggerTrigger:directiondilation - triggerNoTrigger:directiondilation)")

fit.brm_direction_item |> emmeans(pairwise ~ trigger | direction)
```

```{r fig.width=8, fig.height=5, fig.dpi=300}
# Posterior contrast: trigger effect per direction
direction_draws <- as_draws_df(fit.brm_direction_item) |>
  transmute(
    Constriction = `b_triggerTrigger:directionconstriction` -
                   `b_triggerNoTrigger:directionconstriction`,
    Dilation     = `b_triggerTrigger:directiondilation` -
                   `b_triggerNoTrigger:directiondilation`
  ) |>
  pivot_longer(everything(), names_to = "direction", values_to = "trigger_effect") |>
  mutate(direction = factor(direction, levels = c("Constriction", "Dilation")))

plot.direction_trigger <- ggplot(direction_draws,
    aes(x = trigger_effect, y = direction, fill = direction)) +
  stat_halfeye(
    point_interval = median_hdi, .width = c(0.66, 0.9),
    slab_alpha = 0.8, normalize = "panels", point_size = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66) +
  scale_fill_manual(values = c("Constriction" = "#4575B4", "Dilation" = "#D73027")) +
  labs(
    x = "Posterior trigger effect (Trigger \u2013 NoTrigger, item d\u2032)",
    y = NULL,
    title = "Trigger rescue by pupil direction"
  ) +
  theme(legend.position = "none")

plot.direction_trigger
```

## Hidden trigger effect

```{r}
# Filter to the two reclassified cells of interest
df.cell_comparison <- df.trials_reclassified |>
  filter(!trial_burnin) |>
  mutate(
    reclassified_cell = case_when(
      trigger_status_original_4cat == "lapse_notrigger" &
        trigger_status_new == "nolapse_notrigger" ~ "A_lapse_notrig_to_nolapse",
      trigger_status_original_4cat == "lapse_trigger" &
        trigger_status_new == "nolapse_trigger" ~ "B_lapse_trig_to_nolapse",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(reclassified_cell))

message("Trials per reclassified cell:")
df.cell_comparison |> count(reclassified_cell)
```

```{r}
# Item d' for hidden trigger effect
df.cell_item_dprime <- df.cell_comparison |>
  filter(!resp_item_mem_acc %in% c("blink", "NO_RESPONSE")) |>
  group_by(participant_id, reclassified_cell) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_target = sum(goal %in% c("BGSM", "PLUP")),
    n_foil = sum(goal == "NOV"),
    n_trials = n(),
    .groups = "drop"
  ) |>
  mutate(item_dp = fun.dprime(item_hit, item_fa, n_target, n_foil))

df.cell_item_wide <- df.cell_item_dprime |>
  select(participant_id, reclassified_cell, item_dp, n_trials) |>
  pivot_wider(names_from = reclassified_cell, values_from = c(item_dp, n_trials)) |>
  filter(!is.na(item_dp_A_lapse_notrig_to_nolapse),
         !is.na(item_dp_B_lapse_trig_to_nolapse)) |>
  mutate(delta_dp = item_dp_B_lapse_trig_to_nolapse - item_dp_A_lapse_notrig_to_nolapse)

cat("=== ITEM d': Hidden Trigger Effect ===\n")
cat(sprintf("Subjects: %d\n\n", nrow(df.cell_item_wide)))

cat(sprintf("Cell A (lapse_notrig -> nolapse): M=%.3f, SD=%.3f, trials=%.1f\n",
    mean(df.cell_item_wide$item_dp_A_lapse_notrig_to_nolapse),
    sd(df.cell_item_wide$item_dp_A_lapse_notrig_to_nolapse),
    mean(df.cell_item_wide$n_trials_A_lapse_notrig_to_nolapse)))
cat(sprintf("Cell B (lapse_trig -> nolapse):   M=%.3f, SD=%.3f, trials=%.1f\n\n",
    mean(df.cell_item_wide$item_dp_B_lapse_trig_to_nolapse),
    sd(df.cell_item_wide$item_dp_B_lapse_trig_to_nolapse),
    mean(df.cell_item_wide$n_trials_B_lapse_trig_to_nolapse)))

t_hidden_item <- t.test(
  df.cell_item_wide$item_dp_B_lapse_trig_to_nolapse,
  df.cell_item_wide$item_dp_A_lapse_notrig_to_nolapse,
  paired = TRUE, alternative = "greater"
)
print(t_hidden_item)

d_hidden_item <- effectsize::cohens_d(
  df.cell_item_wide$item_dp_B_lapse_trig_to_nolapse,
  df.cell_item_wide$item_dp_A_lapse_notrig_to_nolapse,
  paired = TRUE
)
cat(sprintf("\nCohen's d: %.3f | Mean delta: %.3f\n",
    d_hidden_item$Cohens_d, mean(df.cell_item_wide$delta_dp)))
```

## Hidden trigger effect: Bayesian models

### Item d'

```{r warning=FALSE}
df.cell_item_dprime <- df.cell_item_dprime |>
  mutate(condition = factor(reclassified_cell,
    levels = c("A_lapse_notrig_to_nolapse", "B_lapse_trig_to_nolapse"),
    labels = c("NoTrigger", "Trigger")))

fit.brm_hidden_item = brm(
  item_dp ~ 0 + condition,
  data    = df.cell_item_dprime,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_hidden_trigger_item_dprime"
)

fit.brm_hidden_item |> summary()

# Trigger > NoTrigger (hidden trigger effect)
hypothesis(fit.brm_hidden_item, "conditionTrigger > conditionNoTrigger")

fit.brm_hidden_item |> emmeans(pairwise ~ condition)
```

### Source d'

```{r}
df.cell_source_dprime <- df.cell_comparison |>
  filter(!resp_src_mem_acc %in% c("blink", "NO_RESPONSE")) |>
  group_by(participant_id, reclassified_cell) |>
  summarise(
    src_hit = sum(resp_src_mem_acc %in% c("Size Hit", "Pleasant Hit")),
    src_fa = sum(resp_src_mem_acc %in% c("Size FA", "Pleasant FA")),
    n_src_target = sum(goal %in% c("BGSM", "PLUP")),
    n_src_foil = sum(goal == "NOV"),
    n_trials = n(),
    .groups = "drop"
  ) |>
  mutate(
    src_dp = fun.dprime(src_hit, src_fa, n_src_target, n_src_foil),
    condition = factor(reclassified_cell,
      levels = c("A_lapse_notrig_to_nolapse", "B_lapse_trig_to_nolapse"),
      labels = c("NoTrigger", "Trigger"))
  )
```

```{r warning=FALSE}
fit.brm_hidden_source = brm(
  src_dp ~ 0 + condition,
  data    = df.cell_source_dprime,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_hidden_trigger_source_dprime"
)

fit.brm_hidden_source |> summary()

# Trigger > NoTrigger (hidden trigger effect for source)
hypothesis(fit.brm_hidden_source, "conditionTrigger > conditionNoTrigger")

fit.brm_hidden_source |> emmeans(pairwise ~ condition)
```

## Within-subject correlation: Original vs. hidden trigger effect

```{r}
# Compute original trigger effect per subject (LapseTrigger - Lapse item d')
df.original_trigger_effect <- df.trials_reclassified |>
  filter(!trial_burnin) |>
  filter(trigger_status %in% c("lapse_no_trigger", "lapse_trigger")) |>
  group_by(participant_id, trigger_status) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_item_target = sum(goal %in% c("BGSM", "PLUP")),
    n_item_foil = sum(goal == "NOV"),
    .groups = "drop"
  ) |>
  mutate(item_dp = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil)) |>
  select(participant_id, trigger_status, item_dp) |>
  pivot_wider(names_from = trigger_status, values_from = item_dp) |>
  mutate(original_trigger_effect = lapse_trigger - lapse_no_trigger)

# Merge with hidden trigger effect
df.trigger_correlation <- df.original_trigger_effect |>
  inner_join(
    df.cell_item_wide |> select(participant_id, delta_dp),
    by = "participant_id"
  ) |>
  rename(hidden_trigger_effect = delta_dp)

message(glue("Subjects with both effects: {nrow(df.trigger_correlation)}"))

# Frequentist correlation
cor.test(df.trigger_correlation$original_trigger_effect,
         df.trigger_correlation$hidden_trigger_effect)
```

```{r warning=FALSE}
fit.brm_trigger_correlation = brm(
  hidden_trigger_effect ~ original_trigger_effect,
  data    = df.trigger_correlation,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_trigger_correlation"
)

fit.brm_trigger_correlation |> summary()

# Positive slope (common mechanism)
hypothesis(fit.brm_trigger_correlation, "original_trigger_effect > 0")
```

```{r fig.width=6, fig.height=5, fig.dpi=300}
# Posterior regression lines
post_draws <- as_draws_df(fit.brm_trigger_correlation) |>
  select(b_Intercept, b_original_trigger_effect) |>
  slice_sample(n = 200)

r_val <- cor(df.trigger_correlation$original_trigger_effect,
             df.trigger_correlation$hidden_trigger_effect)

plot.trigger_correlation <- ggplot(df.trigger_correlation,
       aes(x = original_trigger_effect, y = hidden_trigger_effect)) +
  geom_abline(
    data = post_draws,
    aes(intercept = b_Intercept, slope = b_original_trigger_effect),
    alpha = 0.05, color = "steelblue"
  ) +
  geom_point(size = 2.5, alpha = 0.7) +
  geom_abline(
    intercept = fixef(fit.brm_trigger_correlation)[1, 1],
    slope = fixef(fit.brm_trigger_correlation)[2, 1],
    color = "steelblue", linewidth = 1.2
  ) +
  # annotate("text",
  #   x = min(df.trigger_correlation$original_trigger_effect) + 0.1,
  #   y = max(df.trigger_correlation$hidden_trigger_effect) - 0.1,
  #   label = sprintf("r = %.2f", r_val),
  #   size = 5, hjust = 0, vjust = 1, family = "Inter"
  # ) +
  labs(
    x = "Original trigger effect (\u0394 item d\u2032)",
    y = "Hidden trigger effect (\u0394 item d\u2032)"
    # title = "Within-subject correlation: Original vs. hidden trigger"
  )

plot.trigger_correlation
```

## Trial position confound analysis

### Item d' across trial position tertiles

```{r}
# Add trial position tertiles within each run
df.tertile <- df.trials_reclassified |>
  filter(!trial_burnin) |>
  group_by(participant_id, run) |>
  mutate(trial_tertile = ntile(trial, 3)) |>
  ungroup() |>
  mutate(trial_tertile_label = factor(trial_tertile,
    levels = 1:3, labels = c("Early", "Middle", "Late")))

# Item d' per subject per tertile (overall, collapsing across conditions)
df.tertile_item_dp <- df.tertile |>
  group_by(participant_id, trial_tertile_label) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_item_target = sum(goal %in% c("BGSM", "PLUP")),
    n_item_foil = sum(goal == "NOV"),
    .groups = "drop"
  ) |>
  mutate(item_dp = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil))

# Mean d' by tertile
df.tertile_item_dp |>
  group_by(trial_tertile_label) |>
  summarise(
    mean_dp = round(mean(item_dp), 2),
    sd_dp = round(sd(item_dp), 2),
    n = n(),
    .groups = "drop"
  )
```

### Lapse effect within each tertile

```{r}
# Item d' per subject per tertile per lapse condition
df.tertile_lapse_dp <- df.tertile |>
  filter(trigger_status %in% c("no_lapse_no_trigger", "lapse_no_trigger")) |>
  mutate(lapse_condition = ifelse(trigger_status == "no_lapse_no_trigger", "NoLapse", "Lapse")) |>
  group_by(participant_id, trial_tertile_label, lapse_condition) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_item_target = sum(goal %in% c("BGSM", "PLUP")),
    n_item_foil = sum(goal == "NOV"),
    .groups = "drop"
  ) |>
  mutate(item_dp = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil))

# Pivot wide and compute lapse effect per subject per tertile
df.tertile_lapse_wide <- df.tertile_lapse_dp |>
  select(participant_id, trial_tertile_label, lapse_condition, item_dp) |>
  pivot_wider(names_from = lapse_condition, values_from = item_dp) |>
  filter(!is.na(NoLapse), !is.na(Lapse)) |>
  mutate(delta_dp = NoLapse - Lapse)

# Paired t-test within each tertile
df.tertile_lapse_wide |>
  group_by(trial_tertile_label) |>
  summarise(
    mean_delta = round(mean(delta_dp), 2),
    t_stat = round(t.test(delta_dp)$statistic, 2),
    df = t.test(delta_dp)$parameter,
    p_value = round(t.test(delta_dp)$p.value, 3),
    .groups = "drop"
  )
```

### Correlation: Trial position difference vs. d' difference

```{r}
# Per subject: mean trial position for no-lapse vs lapse trials
df.position_diff <- df.tertile |>
  filter(trigger_status %in% c("no_lapse_no_trigger", "lapse_no_trigger")) |>
  mutate(lapse_condition = ifelse(trigger_status == "no_lapse_no_trigger", "NoLapse", "Lapse")) |>
  group_by(participant_id, lapse_condition) |>
  summarise(mean_position = mean(trial), .groups = "drop") |>
  pivot_wider(names_from = lapse_condition, values_from = mean_position) |>
  mutate(position_diff = NoLapse - Lapse)

# Per subject: d' difference (overall, not split by tertile)
df.dprime_diff <- df.tertile |>
  filter(trigger_status %in% c("no_lapse_no_trigger", "lapse_no_trigger")) |>
  mutate(lapse_condition = ifelse(trigger_status == "no_lapse_no_trigger", "NoLapse", "Lapse")) |>
  group_by(participant_id, lapse_condition) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_item_target = sum(goal %in% c("BGSM", "PLUP")),
    n_item_foil = sum(goal == "NOV"),
    .groups = "drop"
  ) |>
  mutate(item_dp = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil)) |>
  select(participant_id, lapse_condition, item_dp) |>
  pivot_wider(names_from = lapse_condition, values_from = item_dp) |>
  mutate(dp_diff = NoLapse - Lapse)

# Merge and correlate
df.position_dp_corr <- df.position_diff |>
  inner_join(df.dprime_diff |> select(participant_id, dp_diff), by = "participant_id")

cor.test(df.position_dp_corr$position_diff, df.position_dp_corr$dp_diff)
```

```{r fig.width=6, fig.height=5, fig.dpi=300}
r_pos <- cor(df.position_dp_corr$position_diff, df.position_dp_corr$dp_diff)

plot.position_confound <- ggplot(df.position_dp_corr,
       aes(x = position_diff, y = dp_diff)) +
  geom_point(size = 2.5, alpha = 0.7) +
  geom_smooth(method = "lm", color = "steelblue", fill = "steelblue", alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  annotate("text",
    x = min(df.position_dp_corr$position_diff) + 0.5,
    y = max(df.position_dp_corr$dp_diff) - 0.1,
    label = sprintf("r = %.2f", r_pos),
    size = 5, hjust = 0, vjust = 1, family = "Inter"
  ) +
  labs(
    x = "Mean trial position difference (NoLapse \u2212 Lapse)",
    y = "Item d\u2032 difference (NoLapse \u2212 Lapse)",
    title = "Trial position vs. memory confound check"
  )

plot.position_confound
```

## Composite figure

```{r fig.width=14, fig.height=12, fig.dpi=300}
# Hidden trigger contrast plot (item + source in one panel)
hidden_item_draws <- as_draws_df(fit.brm_hidden_item) %>%
  transmute(
    contrast = b_conditionTrigger - b_conditionNoTrigger,
    outcome = "Item d\u2032"
  )

hidden_source_draws <- as_draws_df(fit.brm_hidden_source) %>%
  transmute(
    contrast = b_conditionTrigger - b_conditionNoTrigger,
    outcome = "Source d\u2032"
  )

hidden_draws <- bind_rows(hidden_item_draws, hidden_source_draws) %>%
  mutate(outcome = factor(outcome, levels = c("Source d\u2032", "Item d\u2032")))

plot.hidden_trigger_contrasts <- ggplot(hidden_draws, aes(x = contrast, y = outcome, fill = outcome)) +
  stat_halfeye(
    point_interval = median_hdi, .width = c(0.66, 0.9),
    slab_alpha = 0.8, normalize = "panels", point_size = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66, color = "black") +
  scale_fill_manual(values = c("Item d\u2032" = "#F67E00", "Source d\u2032" = "#FCBF4A")) +
  labs(
    x = "Posterior contrast (Trigger \u2013 Lapse)",
    y = NULL,
    title = "Hidden trigger effect"
  ) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

# Compose figure
# plot.spline_spaghetti /
  (plot.spline_item_memory_contrasts | plot.spline_source_memory_contrasts) /
  (plot.hidden_trigger_contrasts | plot.trigger_correlation) +
  plot_annotation(tag_levels = "a") &
  theme(plot.tag = element_text(size = 32, face = "bold", family = "Arial"))

ggsave(here("plots", "fig4.svg"),
       width = 14, height = 12, dpi = 300, bg = "white")
```

## Per-subject spline diagnostic PDF

```{r}
# Build plot data directly from df.comparison
# Shape = original classification, Color = new spline classification
df.plot_data <- df.comparison |>
  mutate(
    # Use window_two assay if it exists, else window_one
    assay_value = coalesce(trigger_assay_two_mean, trigger_assay_one_mean),
    # Per-trial spline trend estimate (for raw-space trend line)
    spline_trend_est = realtime_pupil_raw - realtime_pupil_detrended,
    # Original classification (determines shape)
    orig_class = case_when(
      trial_burnin ~ "burnin",
      trigger_status_original_4cat == "nolapse_notrigger" ~ "nolapse_notrigger",
      trigger_status_original_4cat == "lapse_notrigger"   ~ "lapse_notrigger",
      trigger_status_original_4cat == "lapse_trigger"     ~ "lapse_trigger",
      TRUE ~ NA_character_
    ),
    # New spline classification (determines color)
    new_class = case_when(
      trial_burnin ~ "burnin",
      is.na(trigger_status_new) ~ NA_character_,
      grepl("^nolapse", trigger_status_new) ~ "nolapse",
      grepl("^lapse", trigger_status_new)   ~ "lapse",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(orig_class), !is.na(new_class))

# Shape: original classification
orig_shapes <- c(
  "nolapse_notrigger" = 15,  # filled square
  "lapse_notrigger"   = 17,  # filled triangle
  "lapse_trigger"     = 16,  # filled circle
  "burnin"            = 1    # open circle
)

# Color: new spline classification
new_colors <- c(
  "nolapse" = "#2CA02C",  # green
  "lapse"   = "#D62729",  # red
  "burnin"  = "gray75"
)

df.plot_data <- df.plot_data |>
  mutate(
    orig_class = factor(orig_class, levels = names(orig_shapes)),
    new_class  = factor(new_class, levels = names(new_colors))
  )

message(glue("Plot data: {nrow(df.plot_data)} trials from {n_distinct(df.plot_data$subject_id)} subjects"))
df.plot_data |> count(orig_class, new_class)
```

```{r}
pdf(here("plots", "subject_spline_diagnostics.pdf"), width = 16, height = 16)

for (subj in sort(unique(df.plot_data$subject_id))) {
  df.subj <- df.plot_data |> filter(subject_id == subj)
  df.subj_nonburnin <- df.subj |> filter(!trial_burnin)

  # --- Top panel: raw pupil space ---
  p_raw <- ggplot() +
    # Per-trial spline trend estimate
    geom_line(data = df.subj,
              aes(x = stim_index, y = spline_trend_est),
              color = "red", linewidth = 0.8) +
    # Original thresholds (non-burn-in only — constant after burn-in)
    geom_line(data = df.subj_nonburnin,
              aes(x = stim_index, y = trigger_baseline_mean),
              color = "black", linewidth = 0.5) +
    geom_line(data = df.subj_nonburnin,
              aes(x = stim_index, y = trigger_upper_threshold),
              color = "black", linewidth = 0.4, linetype = "dashed") +
    geom_line(data = df.subj_nonburnin,
              aes(x = stim_index, y = trigger_lower_threshold),
              color = "black", linewidth = 0.4, linetype = "dashed") +
    # Trial assay points
    geom_point(data = df.subj,
               aes(x = stim_index, y = assay_value,
                   color = new_class, shape = orig_class),
               size = 2.5, alpha = 0.85) +
    scale_color_manual(values = new_colors, drop = FALSE) +
    scale_shape_manual(values = orig_shapes, drop = FALSE) +
    facet_wrap(~ run, ncol = 1,
               labeller = labeller(run = function(x) paste("Run", x))) +
    labs(
      title = subj,
      subtitle = "Raw space: red = spline trend, black = original threshold (mean \u00B1 1 SD)",
      x = NULL, y = "Pupil size (raw)",
      shape = "Original", color = "Spline"
    ) +
    theme_classic(base_size = 11) +
    theme(legend.position = "none",
          strip.text = element_text(size = 11, face = "bold"))

  # --- Bottom panel: detrended space ---
  df.run_thresholds <- df.subj |>
    group_by(run) |>
    summarise(
      global_upper = first(global_upper_threshold),
      global_lower = first(global_lower_threshold),
      global_mean  = first(global_realtime_mean),
      .groups = "drop"
    )

  p_detrended <- ggplot() +
    # Flat spline thresholds (constant per run)
    geom_hline(data = df.run_thresholds,
               aes(yintercept = global_mean),
               color = "steelblue", linewidth = 0.5) +
    geom_hline(data = df.run_thresholds,
               aes(yintercept = global_upper),
               color = "steelblue", linewidth = 0.4, linetype = "dashed") +
    geom_hline(data = df.run_thresholds,
               aes(yintercept = global_lower),
               color = "steelblue", linewidth = 0.4, linetype = "dashed") +
    # Detrended trial points (the actual values used for classification)
    geom_point(data = df.subj,
               aes(x = stim_index, y = realtime_pupil_detrended,
                   color = new_class, shape = orig_class),
               size = 2.5, alpha = 0.85) +
    scale_color_manual(values = new_colors, drop = FALSE) +
    scale_shape_manual(values = orig_shapes, drop = FALSE) +
    facet_wrap(~ run, ncol = 1,
               labeller = labeller(run = function(x) paste("Run", x))) +
    labs(
      subtitle = "Detrended space: blue = global spline threshold (mean \u00B1 1 SD)",
      x = "Stimulus index", y = "Pupil size (detrended)",
      shape = "Original", color = "Spline"
    ) +
    theme_classic(base_size = 11) +
    theme(
      legend.position = "bottom",
      legend.title = element_text(size = 10),
      legend.text = element_text(size = 9),
      strip.text = element_text(size = 11, face = "bold")
    ) +
    guides(
      shape = guide_legend(order = 1, override.aes = list(size = 3)),
      color = guide_legend(order = 2, override.aes = list(size = 3))
    )

  print(p_raw / p_detrended + plot_layout(heights = c(1, 1)))
}

dev.off()
message("Saved: plots/subject_spline_diagnostics.pdf")
```

