# Spline detrending v2 (continuous timeseries)

## Parameters
```{r}
SPLINE_DF = 5
OUTLIER_THRESHOLD = 1.0

PUPIL_COL = "pupil_raw_deblink_detransient_interpolate_lpfilt"

# Path to eyeris RDS files
path.eyeris_dir = "/Volumes/LumpySpace/clamp/derivatives"
path.spline_cache = here("models", "splines")
```

## Load trial-level behavioral data
```{r}
# Full trial order (before any filtering)
# Some stimuli repeat within a run — prefer non-aborted occurrence
df.trial_order <- df.trials |>
  select(participant_id, run, stim, stim_index, trial, trial_aborted) |>
  group_by(participant_id, run, stim) |>
  arrange(trial_aborted, desc(stim_index)) |>
  slice(1) |>
  ungroup() |>
  select(-trial_aborted)

df.trials_v2 <- df.trials |>
  filter(!participant_id %in% ids_to_exclude) |>
  filter(!resp_item_mem_acc %in% c("blink", "NO_RESPONSE"))

message(paste("Loaded", nrow(df.trials_v2), "trials"))
```

## Spline fitting and caching
```{r}
fit_and_cache_spline = function(subj_id, run_num, spline_df = SPLINE_DF,
                                 eyeris_dir = path.eyeris_dir,
                                 cache_dir = path.spline_cache,
                                 pupil_col = PUPIL_COL) {
  # Manual exclusions
  if (subj_id == "sub-71" && run_num == 3) return(NULL)

  # Cache path: models/splines/sub-01/run-01_spline.rds
  subj_cache_dir <- file.path(cache_dir, subj_id)
  cache_file <- file.path(subj_cache_dir, paste0("run-", str_pad(run_num, 2, pad = "0"), "_spline.rds"))

  if (file.exists(cache_file)) {
    return(readRDS(cache_file))
  }

  # Load RDS
  rds_path <- file.path(eyeris_dir, subj_id, "ses-ret", "eye",
    paste0(subj_id, "_ses-ret_task-clamp_run-", str_pad(run_num, 2, pad = "0"),
           "_desc-eyeris_preproc_epoched.RDS"))

  if (!file.exists(rds_path)) {
    warning(paste("RDS not found:", rds_path))
    return(NULL)
  }

  rds <- readRDS(rds_path)
  block_name <- paste0("block_", run_num)
  ts <- rds$timeseries[[block_name]]

  if (is.null(ts) || nrow(ts) == 0) {
    warning(paste("Empty timeseries for", subj_id, "run", run_num))
    return(NULL)
  }

  # Fit spline to full continuous timeseries
  spline_basis <- splines::ns(ts$time_secs, df = spline_df)
  model <- lm(ts[[pupil_col]] ~ spline_basis)
  ts$spline_trend <- fitted(model)
  ts$pupil_detrended <- ts[[pupil_col]] - ts$spline_trend

  # Extract spline parameters for predicting at arbitrary timepoints
  spline_params <- list(
    knots = attr(spline_basis, "knots"),
    boundary_knots = attr(spline_basis, "Boundary.knots"),
    coefficients = coef(model)
  )

  # Save the spline params + the timeseries with trend
  result <- list(
    subject_id = subj_id,
    run = run_num,
    spline_params = spline_params,
    timeseries = ts |>
      select(time_secs, all_of(pupil_col), spline_trend, pupil_detrended) |>
      filter(row_number() %% 10 == 1),
    epochs = list()
  )

  # Extract per-trial realtime values from epoched data
  ep_rt1 <- rds$epoch_realtime1[[block_name]]
  if (is.null(ep_rt1) || nrow(ep_rt1) == 0) {
    warning(paste("No realtime1 epochs for", subj_id, "run", run_num))
    return(NULL)
  }
  ep_rt1 <- ep_rt1 |>
    mutate(epoch = "realtime1") |>
    select(time_secs, timebin, STIM, epoch, pupil = all_of(pupil_col))

  # realtime2: only trials that went to second window
  ep_rt2 <- rds$epoch_realtime2[[block_name]]
  if (!is.null(ep_rt2) && nrow(ep_rt2) > 0) {
    ep_rt2 <- ep_rt2 |>
      mutate(epoch = "realtime2") |>
      select(time_secs, timebin, STIM, epoch, pupil = all_of(pupil_col))
  } else {
    ep_rt2 <- tibble(time_secs = numeric(), timebin = numeric(),
                     STIM = character(), epoch = character(), pupil = numeric())
  }

  ep_all <- bind_rows(ep_rt1, ep_rt2)

  # Look up spline trend at each epoch timepoint
  ep_basis <- splines::ns(ep_all$time_secs,
    knots = spline_params$knots,
    Boundary.knots = spline_params$boundary_knots
  )
  ep_all$spline_trend <- as.numeric(cbind(1, ep_basis) %*% spline_params$coefficients)

  # Detrend
  ep_all$pupil_detrended <- ep_all$pupil - ep_all$spline_trend

  # Summarize per trial: mean detrended value across the relevant realtime window(s)
  trial_summary <- ep_all |>
    group_by(STIM) |>
    summarise(
      realtime_pupil_raw = mean(pupil, na.rm = TRUE),
      realtime_pupil_detrended = mean(pupil_detrended, na.rm = TRUE),
      spline_trend_at_rt = mean(spline_trend, na.rm = TRUE),
      rt_time_secs = mean(time_secs),
      n_rt_samples = n(),
      .groups = "drop"
    ) |>
    rename(item = STIM)

  result$trial_summary <- trial_summary
  result$epoch_data <- ep_all

  # Cache
  dir.create(subj_cache_dir, recursive = TRUE, showWarnings = FALSE)
  saveRDS(result, cache_file)

  return(result)
}
```

## Process single subject (all runs)
```{r}
process_subject_v2 = function(subj_id, runs = 1:3, outlier_threshold = OUTLIER_THRESHOLD) {

  run_results <- list()
  for (r in runs) {
    res <- fit_and_cache_spline(subj_id, r)
    if (!is.null(res)) run_results[[length(run_results) + 1]] <- res
  }

  if (length(run_results) == 0) return(NULL)

  # Combine trial summaries across runs
  df.trial_summary <- map_dfr(run_results, function(res) {
    res$trial_summary |> mutate(subject_id = res$subject_id, run = res$run)
  })

  # Join with behavioral data to get original classification + stim_index
  df.trial_summary <- df.trial_summary |>
    left_join(
      df.trial_order |>
        rename(subject_id = participant_id, item = stim),
      by = c("subject_id", "run", "item")
    )

  # Compute global stats per run (excluding burn-in)
  # Use df.trial_order source (df.trials) for burn-in info — join via stim_index
  df.trial_summary <- df.trial_summary |>
    left_join(
      df.trials |>
        select(participant_id, run, stim, stim_index, trial_aborted, trial_burnin) |>
        group_by(participant_id, run, stim) |>
        arrange(trial_aborted, desc(stim_index)) |>
        slice(1) |>
        ungroup() |>
        rename(subject_id = participant_id, item = stim) |>
        select(-stim_index, -trial_aborted),
      by = c("subject_id", "run", "item")
    )

  df.global_stats <- df.trial_summary |>
    filter(!trial_burnin | is.na(trial_burnin)) |>
    group_by(subject_id, run) |>
    summarise(
      global_realtime_mean = mean(realtime_pupil_detrended, na.rm = TRUE),
      global_realtime_sd = sd(realtime_pupil_detrended, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(
      global_upper_threshold = global_realtime_mean + outlier_threshold * global_realtime_sd,
      global_lower_threshold = global_realtime_mean - outlier_threshold * global_realtime_sd
    )

  # Classify
  df.trial_summary <- df.trial_summary |>
    left_join(df.global_stats, by = c("subject_id", "run")) |>
    mutate(
      is_lapse_spline = case_when(
        trial_burnin ~ NA,
        realtime_pupil_detrended > global_upper_threshold |
          realtime_pupil_detrended < global_lower_threshold ~ TRUE,
        TRUE ~ FALSE
      )
    )

  # Join with df.trials to get original thresholds and classification
  # Keep non-aborted occurrence per stim (prefer completed trials over aborted ones)
  df.merged <- df.trial_summary |>
    left_join(
      df.trials |>
        select(participant_id, run, stim, stim_index, trial_aborted, trigger_status,
               trigger_baseline_mean, trigger_baseline_std,
               trigger_upper_threshold, trigger_lower_threshold,
               trigger_assay_one_mean, trigger_assay_two_mean,
               trigger_actually_delivered, realtime_window_status,
               window_one_status, window_two_status,
               latest_baseline_mean) |>
        group_by(participant_id, run, stim) |>
        arrange(trial_aborted, desc(stim_index)) |>
        slice(1) |>
        ungroup() |>
        select(-stim_index, -trial_aborted) |>
        rename(subject_id = participant_id, item = stim),
      by = c("subject_id", "run", "item")
    ) |>
    mutate(
      assay_value = case_when(
        trial_burnin ~ latest_baseline_mean,
        TRUE ~ coalesce(trigger_assay_two_mean, trigger_assay_one_mean, realtime_pupil_raw)
      ),
      trigger_status_new = case_when(
        is.na(is_lapse_spline) ~ NA_character_,
        is_lapse_spline & trigger_actually_delivered ~ "lapse_trigger",
        is_lapse_spline & !trigger_actually_delivered ~ "lapse_notrigger",
        !is_lapse_spline & !trigger_actually_delivered ~ "nolapse_notrigger",
        !is_lapse_spline & trigger_actually_delivered ~ "nolapse_trigger"
      ),
      # Original 4-category status (standardized)
      trigger_status_original_4cat = case_when(
        trigger_status == "lapse_trigger" ~ "lapse_trigger",
        trigger_status == "lapse_no_trigger" ~ "lapse_notrigger",
        trigger_status == "no_lapse_no_trigger" ~ "nolapse_notrigger",
        TRUE ~ trigger_status
      ),
      # Plot aesthetics
      orig_class = case_when(
        trial_burnin ~ "burnin",
        trigger_status_original_4cat == "nolapse_notrigger" ~ "nolapse_notrigger",
        trigger_status_original_4cat == "lapse_notrigger"   ~ "lapse_notrigger",
        trigger_status_original_4cat == "lapse_trigger"     ~ "lapse_trigger",
        TRUE ~ NA_character_
      ),
      new_class = case_when(
        trial_burnin ~ "burnin",
        is.na(trigger_status_new) ~ NA_character_,
        grepl("^nolapse", trigger_status_new) ~ "nolapse",
        grepl("^lapse", trigger_status_new)   ~ "lapse",
        TRUE ~ NA_character_
      )
    )

  list(
    trial_data = df.merged,
    run_results = run_results
  )
}
```

## Test on sub-01
```{r}
result_sub01 <- process_subject_v2("sub-01")

message(glue("Trials: {nrow(result_sub01$trial_data)}"))
result_sub01$trial_data |> count(orig_class, new_class)
```

## Test on sub-05
```{r}
result_sub05 <- process_subject_v2("sub-05")

message(glue("Trials: {nrow(result_sub05$trial_data)}"))
result_sub05$trial_data |> count(orig_class, new_class)
```

```{r fig.width=16, fig.height=16}
plot_subject_diagnostic(result_sub05, "sub-05")
```

## Diagnostic plot for one subject
```{r fig.width=16, fig.height=16}
# Visual encoding
orig_shapes <- c(
  "nolapse_notrigger" = 15,  # filled square
  "lapse_notrigger"   = 17,  # filled triangle
  "lapse_trigger"     = 16,  # filled circle
  "burnin"            = 1    # open circle
)

new_colors <- c(
  "nolapse" = "#2CA02C",
  "lapse"   = "#D62729",
  "burnin"  = "gray75"
)

plot_subject_diagnostic = function(result, subj_id) {
  df <- result$trial_data |>
    filter(!is.na(orig_class), !is.na(new_class)) |>
    mutate(
      orig_class = factor(orig_class, levels = names(orig_shapes)),
      new_class  = factor(new_class, levels = names(new_colors))
    )

  df_nonburnin <- df |> filter(!trial_burnin)

  # Spline curve from cached timeseries (already downsampled)
  df_spline_curve <- map_dfr(result$run_results, function(res) {
    res$timeseries |> mutate(run = res$run)
  })

  # --- Top panel: raw pupil space ---
  p_raw <- ggplot() +
    # Raw continuous pupil trace (background)
    geom_line(data = df_spline_curve,
              aes(x = time_secs, y = pupil_raw_deblink_detransient_interpolate_lpfilt),
              color = "gray80", linewidth = 0.3, alpha = 0.5) +
    # Spline trend
    geom_line(data = df_spline_curve,
              aes(x = time_secs, y = spline_trend),
              color = "red", linewidth = 0.8) +
    geom_line(data = df_nonburnin,
              aes(x = rt_time_secs, y = trigger_baseline_mean),
              color = "black", linewidth = 0.5) +
    geom_line(data = df_nonburnin,
              aes(x = rt_time_secs, y = trigger_upper_threshold),
              color = "black", linewidth = 0.4, linetype = "dashed") +
    geom_line(data = df_nonburnin,
              aes(x = rt_time_secs, y = trigger_lower_threshold),
              color = "black", linewidth = 0.4, linetype = "dashed") +
    geom_point(data = df,
               aes(x = rt_time_secs, y = assay_value,
                   color = new_class, shape = orig_class),
               size = 2.5, alpha = 0.85) +
    scale_color_manual(values = new_colors, drop = FALSE) +
    scale_shape_manual(values = orig_shapes, drop = FALSE) +
    facet_wrap(~ run, ncol = 1, scales = "free_x",
               labeller = labeller(run = function(x) paste("Run", x))) +
    labs(
      title = subj_id,
      subtitle = "Raw space: red = spline trend, black = original threshold (mean \u00B1 1 SD)",
      x = NULL, y = "Pupil size (raw)",
      shape = "Original", color = "Spline"
    ) +
    theme_classic(base_size = 11) +
    theme(legend.position = "none",
          strip.text = element_text(size = 11, face = "bold"))

  # --- Bottom panel: detrended space ---
  df_run_thresholds <- df |>
    group_by(run) |>
    summarise(
      global_upper = first(global_upper_threshold),
      global_lower = first(global_lower_threshold),
      global_mean  = first(global_realtime_mean),
      .groups = "drop"
    )

  p_detrended <- ggplot() +
    geom_hline(data = df_run_thresholds,
               aes(yintercept = global_mean),
               color = "steelblue", linewidth = 0.5) +
    geom_hline(data = df_run_thresholds,
               aes(yintercept = global_upper),
               color = "steelblue", linewidth = 0.4, linetype = "dashed") +
    geom_hline(data = df_run_thresholds,
               aes(yintercept = global_lower),
               color = "steelblue", linewidth = 0.4, linetype = "dashed") +
    geom_point(data = df,
               aes(x = rt_time_secs, y = realtime_pupil_detrended,
                   color = new_class, shape = orig_class),
               size = 2.5, alpha = 0.85) +
    scale_color_manual(values = new_colors, drop = FALSE) +
    scale_shape_manual(values = orig_shapes, drop = FALSE) +
    facet_wrap(~ run, ncol = 1, scales = "free_x",
               labeller = labeller(run = function(x) paste("Run", x))) +
    labs(
      subtitle = "Detrended space: blue = global spline threshold (mean \u00B1 1 SD)",
      x = "Time (s)", y = "Pupil size (detrended)",
      shape = "Original", color = "Spline"
    ) +
    theme_classic(base_size = 11) +
    theme(
      legend.position = "bottom",
      legend.title = element_text(size = 10),
      legend.text = element_text(size = 9),
      strip.text = element_text(size = 11, face = "bold")
    ) +
    guides(
      shape = guide_legend(order = 1, override.aes = list(size = 3)),
      color = guide_legend(order = 2, override.aes = list(size = 3))
    )

  p_raw / p_detrended + plot_layout(heights = c(1, 1))
}

plot_subject_diagnostic(result_sub01, "sub-01")
```

## Generate PDF for all subjects
```{r}
# Get all subject IDs from the eyeris directory
all_subjects <- list.dirs(path.eyeris_dir, full.names = FALSE, recursive = FALSE) |>
  str_subset("^sub-[0-9]+$") |>
  sort()

# Remove excluded subjects
all_subjects <- all_subjects[!all_subjects %in% ids_to_exclude]

message(glue("Processing {length(all_subjects)} subjects for diagnostic PDF..."))

pdf(here("plots", "subject_spline_v2_diagnostics.pdf"), width = 16, height = 16)

for (i in seq_along(all_subjects)) {
  subj <- all_subjects[i]
  if (i %% 5 == 0 || i == 1) message(glue("[{i}/{length(all_subjects)}] {subj}"))

  result <- tryCatch(
    process_subject_v2(subj),
    error = function(e) {
      message(glue("  ERROR {subj}: {e$message}"))
      return(NULL)
    }
  )

  if (is.null(result)) next

  p <- tryCatch(
    plot_subject_diagnostic(result, subj),
    error = function(e) {
      message(glue("  PLOT ERROR {subj}: {e$message}"))
      return(NULL)
    }
  )

  if (!is.null(p)) print(p)

  rm(result, p)
  gc(verbose = FALSE)
}

dev.off()
message("Saved: plots/subject_spline_v2_diagnostics.pdf")
```

## Process all subjects and build reclassification table
```{r}
all_results <- list()

for (subj in all_subjects) {
  res <- tryCatch(
    process_subject_v2(subj),
    error = function(e) {
      message(glue("  ERROR {subj}: {e$message}"))
      return(NULL)
    }
  )
  if (!is.null(res)) all_results[[subj]] <- res
  gc(verbose = FALSE)
}

# Combine all trial data into one dataframe (equivalent of df.trials_reclassified)
df.v2_reclassified <- map_dfr(all_results, ~ .x$trial_data)

# Rename subject_id -> participant_id for consistency with behavioral data joins
df.v2_reclassified <- df.v2_reclassified |>
  rename(participant_id = subject_id)

message(glue("Subjects: {n_distinct(df.v2_reclassified$participant_id)} | Trials: {nrow(df.v2_reclassified)}"))
message(glue("With classification: {sum(!is.na(df.v2_reclassified$trigger_status_new))}"))
```

## Trial counts
```{r}
# Reclassification flow table: Original → Spline v2
message("=== Trial Flow: Original \u2192 Spline v2 ===\n")

flow_table <- df.v2_reclassified |>
  filter(!trial_burnin, !is.na(trigger_status_new)) |>
  count(trigger_status_original_4cat, trigger_status_new) |>
  pivot_wider(names_from = trigger_status_new, values_from = n, values_fill = 0)

flow_table
```

```{r}
# Reclassification flow table split by constriction / dilation
message("=== Trial Flow by Constriction / Dilation ===\n")

flow_by_direction <- df.v2_reclassified |>
  filter(!trial_burnin, !is.na(trigger_status_new),
         realtime_window_status %in% c("constriction", "dilation")) |>
  count(trigger_status_original_4cat, trigger_status_new, realtime_window_status) |>
  pivot_wider(
    names_from = realtime_window_status,
    values_from = n,
    values_fill = 0
  ) |>
  mutate(cell_label = paste0(constriction, " / ", dilation)) |>
  select(trigger_status_original_4cat, trigger_status_new, cell_label) |>
  pivot_wider(
    names_from = trigger_status_new,
    values_from = cell_label,
    values_fill = "0 / 0"
  )

flow_by_direction
```

```{r}
df.v2_reclassified |>
  filter(!trial_burnin, !is.na(trigger_status_original_4cat)) |>
  count(trigger_status_original_4cat) |>
  mutate(pct = round(100 * n / sum(n), 1))

# Trial distribution by spline v2 classification
df.v2_reclassified |>
  filter(!trial_burnin, !is.na(trigger_status_new)) |>
  count(trigger_status_new) |>
  mutate(pct = round(100 * n / sum(n), 1))
```

```{r}
# Per-subject trial counts per condition (summary stats)
message("=== Per-Subject Trial Counts ===\n")

df.v2_reclassified |>
  filter(!is.na(trigger_status_new), !trial_burnin) |>
  count(participant_id, trigger_status_new) |>
  group_by(trigger_status_new) |>
  summarise(
    n_subjects = n(),
    min_trials = min(n),
    median_trials = median(n),
    mean_trials = round(mean(n), 1),
    max_trials = max(n),
    .groups = "drop"
  )
```

## Compute d' per reclassified condition

```{r}
spline_cond_levels = c("nolapse_notrigger", "lapse_notrigger", "lapse_trigger")
spline_cond_labels = c("NoLapse", "Lapse", "LapseTrigger")

# Need to merge behavioral response data for d' computation
df.v2_for_dprime <- df.v2_reclassified |>
  left_join(
    df.trials |>
      select(participant_id, run, stim, stim_index, trial_aborted,
             resp_item_mem_acc, resp_src_mem_acc, goal, stim_old_new) |>
      group_by(participant_id, run, stim) |>
      arrange(trial_aborted, desc(stim_index)) |>
      slice(1) |>
      ungroup() |>
      select(-stim_index, -trial_aborted) |>
      rename(item = stim),
    by = c("participant_id", "run", "item")
  ) |>
  filter(!resp_item_mem_acc %in% c("blink", "NO_RESPONSE"))

# Item d' by subject x reclassified condition
df.spline_item_dp <- df.v2_for_dprime |>
  filter(!trial_burnin, !is.na(trigger_status_new)) |>
  filter(trigger_status_new %in% spline_cond_levels) |>
  group_by(participant_id, trigger_status_new) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_item_target = sum(goal %in% c("BGSM", "PLUP")),
    n_item_foil = sum(goal == "NOV"),
    .groups = "drop"
  ) |>
  mutate(
    item_dp = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil),
    condition = factor(trigger_status_new, levels = spline_cond_levels, labels = spline_cond_labels)
  )

# Source d' by subject x reclassified condition
df.spline_source_dp <- df.v2_for_dprime |>
  filter(!trial_burnin, !is.na(trigger_status_new)) |>
  filter(trigger_status_new %in% spline_cond_levels) |>
  filter(!resp_src_mem_acc %in% c("blink", "NO_RESPONSE")) |>
  group_by(participant_id, trigger_status_new) |>
  summarise(
    src_hit = sum(resp_src_mem_acc %in% c("Size Hit", "Pleasant Hit")),
    src_fa = sum(resp_src_mem_acc %in% c("Size FA", "Pleasant FA")),
    n_src_target = sum(goal %in% c("BGSM", "PLUP")),
    n_src_foil = sum(goal == "NOV"),
    .groups = "drop"
  ) |>
  mutate(
    src_dp = fun.dprime(src_hit, src_fa, n_src_target, n_src_foil),
    condition = factor(trigger_status_new, levels = spline_cond_levels, labels = spline_cond_labels)
  )

message(glue("Item d': {nrow(df.spline_item_dp)} rows from {n_distinct(df.spline_item_dp$participant_id)} subjects"))
message(glue("Source d': {nrow(df.spline_source_dp)} rows from {n_distinct(df.spline_source_dp$participant_id)} subjects"))
```

## Bayesian model: Item d' (spline v2 reclassified)

```{r fig.height=20, warning=FALSE}
fit.brm_spline_item_dprime = brm(
  item_dp ~ 0 + condition,
  data    = df.spline_item_dp,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_v2_spline_item_dprime"
)

fit.brm_spline_item_dprime |> summary()
fit.brm_spline_item_dprime |> plot(nvariables = 8)
fit.brm_spline_item_dprime |> pp_check(ndraws = 100)

fit.brm_spline_item_dprime |>
  as_draws_df() |>
  select(starts_with("b_"), sigma) |>
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") |>
  ggplot(mapping = aes(y = fct_rev(variable), x = value)) +
  stat_halfeye(fill = "lightblue") +
  geom_vline(xintercept = 0.0, linetype = "dashed") +
  theme(axis.title.y = element_blank())

# Lapse < NoLapse
hypothesis(fit.brm_spline_item_dprime, "conditionLapse < conditionNoLapse")

# LapseTrigger > Lapse
hypothesis(fit.brm_spline_item_dprime, "conditionLapseTrigger > conditionLapse")

# LapseTrigger < NoLapse
hypothesis(fit.brm_spline_item_dprime, "conditionLapseTrigger < conditionNoLapse")

fit.brm_spline_item_dprime |>
  emmeans(pairwise ~ condition)
```

```{r fig.width=6, fig.height=5, fig.dpi=300}
contrast_draws <- fit.brm_spline_item_dprime %>%
  spread_draws(b_conditionNoLapse, b_conditionLapse, b_conditionLapseTrigger) %>%
  mutate(
    `Lapse – No Lapse`    = b_conditionLapse - b_conditionNoLapse,
    `Trigger – Lapse`     = b_conditionLapseTrigger - b_conditionLapse,
    `No Lapse – Trigger`  = b_conditionNoLapse - b_conditionLapseTrigger
  ) %>%
  pivot_longer(cols = matches("–"), names_to = "contrast", values_to = ".value") %>%
  mutate(contrast = factor(contrast, levels = c(
    "No Lapse – Trigger", "Trigger – Lapse", "Lapse – No Lapse"
  )))

fill_by_type <- c(
  "Lapse – No Lapse"   = "#D62729",
  "Trigger – Lapse"    = "#F67E00",
  "No Lapse – Trigger" = "#FCBF4A"
)

plot.spline_item_memory_contrasts = contrast_draws %>%
  ggplot(aes(x = .value, y = contrast, fill = contrast)) +
  stat_halfeye(point_interval = median_hdi, .width = c(0.66, 0.9),
               slab_alpha = 0.8, normalize = "panels", point_size = 4) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66, color = "black") +
  geom_text(
    data = contrast_draws %>% group_by(contrast) %>% summarise(m = median(.value)),
    aes(x = m, y = contrast, label = contrast),
    vjust = -0.8, hjust = 0.5, size = 6, inherit.aes = FALSE, family = "Inter"
  ) +
  scale_fill_manual(values = fill_by_type) +
  scale_y_discrete(labels = NULL) +
  labs(x = "Posterior contrast (d\u2032)", y = NULL, title = "Item memory (spline v2)") +
  theme(
    legend.position = "none", panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(), axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

plot.spline_item_memory_contrasts
```

## Bayesian model: Source d' (spline v2 reclassified)

```{r fig.height=20, warning=FALSE}
fit.brm_spline_source_dprime = brm(
  src_dp ~ 0 + condition,
  data    = df.spline_source_dp,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_v2_spline_source_dprime"
)

fit.brm_spline_source_dprime |> summary()
fit.brm_spline_source_dprime |> plot(nvariables = 8)
fit.brm_spline_source_dprime |> pp_check(ndraws = 100)

fit.brm_spline_source_dprime |>
  as_draws_df() |>
  select(starts_with("b_"), sigma) |>
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") |>
  ggplot(mapping = aes(y = fct_rev(variable), x = value)) +
  stat_halfeye(fill = "lightblue") +
  geom_vline(xintercept = 0.0, linetype = "dashed") +
  theme(axis.title.y = element_blank())

# Lapse < NoLapse
hypothesis(fit.brm_spline_source_dprime, "conditionLapse < conditionNoLapse")

# LapseTrigger > Lapse
hypothesis(fit.brm_spline_source_dprime, "conditionLapseTrigger > conditionLapse")

# LapseTrigger < NoLapse
hypothesis(fit.brm_spline_source_dprime, "conditionLapseTrigger < conditionNoLapse")

fit.brm_spline_source_dprime |>
  emmeans(pairwise ~ condition)
```

```{r fig.width=6, fig.height=5, fig.dpi=300}
contrast_draws <- fit.brm_spline_source_dprime %>%
  spread_draws(b_conditionNoLapse, b_conditionLapse, b_conditionLapseTrigger) %>%
  mutate(
    `Lapse – No Lapse`    = b_conditionLapse - b_conditionNoLapse,
    `Trigger – Lapse`     = b_conditionLapseTrigger - b_conditionLapse,
    `No Lapse – Trigger`  = b_conditionNoLapse - b_conditionLapseTrigger
  ) %>%
  pivot_longer(cols = matches("–"), names_to = "contrast", values_to = ".value") %>%
  mutate(contrast = factor(contrast, levels = c(
    "No Lapse – Trigger", "Trigger – Lapse", "Lapse – No Lapse"
  )))

fill_by_type <- c(
  "Lapse – No Lapse"   = "#D62729",
  "Trigger – Lapse"    = "#F67E00",
  "No Lapse – Trigger" = "#FCBF4A"
)

plot.spline_source_memory_contrasts = contrast_draws %>%
  ggplot(aes(x = .value, y = contrast, fill = contrast)) +
  stat_halfeye(point_interval = median_hdi, .width = c(0.66, 0.9),
               slab_alpha = 0.8, normalize = "panels", point_size = 4) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66, color = "black") +
  geom_text(
    data = contrast_draws %>% group_by(contrast) %>% summarise(m = median(.value)),
    aes(x = m, y = contrast, label = contrast),
    vjust = -0.8, hjust = 0.5, size = 6, inherit.aes = FALSE, family = "Inter"
  ) +
  scale_fill_manual(values = fill_by_type) +
  scale_y_discrete(labels = NULL) +
  labs(x = "Posterior contrast (d\u2032)", y = NULL, title = "Source memory (spline v2)") +
  theme(
    legend.position = "none", panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(), axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

plot.spline_source_memory_contrasts
```

## "True lapses" trigger effect

```{r}
# Trials classified as lapse under BOTH original and spline v2 methods
df.true_lapses <- df.v2_for_dprime |>
  filter(!trial_burnin, !is.na(trigger_status_new)) |>
  filter(
    trigger_status_original_4cat %in% c("lapse_trigger", "lapse_notrigger"),
    trigger_status_new %in% c("lapse_trigger", "lapse_notrigger")
  )

true_lapse_cond_levels = c("lapse_notrigger", "lapse_trigger")
true_lapse_cond_labels = c("Lapse", "LapseTrigger")

# Item d' for true lapses
df.true_lapse_item_dp <- df.true_lapses |>
  group_by(participant_id, trigger_status_new) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_item_target = sum(goal %in% c("BGSM", "PLUP")),
    n_item_foil = sum(goal == "NOV"),
    n_trials = n(),
    .groups = "drop"
  ) |>
  mutate(
    item_dp = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil),
    condition = factor(trigger_status_new, levels = true_lapse_cond_levels, labels = true_lapse_cond_labels)
  )

message(glue("True lapses - Item d': {nrow(df.true_lapse_item_dp)} rows from {n_distinct(df.true_lapse_item_dp$participant_id)} subjects"))
df.true_lapse_item_dp |> count(condition)
```

```{r warning=FALSE}
fit.brm_true_lapse_item = brm(
  item_dp ~ 0 + condition,
  data    = df.true_lapse_item_dp,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_v2_true_lapse_item_dprime"
)

fit.brm_true_lapse_item |> summary()

# LapseTrigger > Lapse (trigger rescue among true lapses)
hypothesis(fit.brm_true_lapse_item, "conditionLapseTrigger > conditionLapse")

fit.brm_true_lapse_item |> emmeans(pairwise ~ condition)
```

## Asymmetric trigger effect: Constriction vs. Dilation

```{r}
# Split true lapses by pupil direction (constriction = low arousal, dilation = high arousal)
df.true_lapses_by_direction <- df.true_lapses |>
  filter(realtime_window_status %in% c("constriction", "dilation"))

message(glue("True lapses with direction: {nrow(df.true_lapses_by_direction)} / {nrow(df.true_lapses)} total"))
df.true_lapses_by_direction |> count(trigger_status_new, realtime_window_status)
```

```{r}
# Item d' per subject x trigger condition x pupil direction
df.direction_item_dp <- df.true_lapses_by_direction |>
  group_by(participant_id, trigger_status_new, realtime_window_status) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_item_target = sum(goal %in% c("BGSM", "PLUP")),
    n_item_foil = sum(goal == "NOV"),
    n_trials = n(),
    .groups = "drop"
  ) |>
  mutate(
    item_dp = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil),
    trigger = factor(
      trigger_status_new,
      levels = c("lapse_notrigger", "lapse_trigger"),
      labels = c("NoTrigger", "Trigger")
    ),
    direction = factor(realtime_window_status, levels = c("constriction", "dilation"))
  )

message(glue("Direction x Trigger: {nrow(df.direction_item_dp)} rows from {n_distinct(df.direction_item_dp$participant_id)} subjects"))
df.direction_item_dp |> count(trigger, direction)
```

```{r warning=FALSE}
# 2x2 model: trigger x pupil direction
fit.brm_direction_item = brm(
  item_dp ~ 0 + trigger:direction,
  data    = df.direction_item_dp,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_v2_direction_trigger_item_dprime"
)

fit.brm_direction_item |> summary()

# Trigger effect within constriction trials
hypothesis(fit.brm_direction_item,
  "triggerTrigger:directionconstriction > triggerNoTrigger:directionconstriction")

# Trigger effect within dilation trials
hypothesis(fit.brm_direction_item,
  "triggerTrigger:directiondilation > triggerNoTrigger:directiondilation")

# Asymmetry: trigger effect on constriction > trigger effect on dilation
hypothesis(fit.brm_direction_item,
  "(triggerTrigger:directionconstriction - triggerNoTrigger:directionconstriction) > (triggerTrigger:directiondilation - triggerNoTrigger:directiondilation)")

fit.brm_direction_item |> emmeans(pairwise ~ trigger | direction)
```

```{r fig.width=8, fig.height=5, fig.dpi=300}
# Posterior contrast: trigger effect per direction
direction_draws <- as_draws_df(fit.brm_direction_item) |>
  transmute(
    Constriction = `b_triggerTrigger:directionconstriction` -
                   `b_triggerNoTrigger:directionconstriction`,
    Dilation     = `b_triggerTrigger:directiondilation` -
                   `b_triggerNoTrigger:directiondilation`
  ) |>
  pivot_longer(everything(), names_to = "direction", values_to = "trigger_effect") |>
  mutate(direction = factor(direction, levels = c("Constriction", "Dilation")))

plot.direction_trigger <- ggplot(direction_draws,
    aes(x = trigger_effect, y = direction, fill = direction)) +
  stat_halfeye(
    point_interval = median_hdi, .width = c(0.66, 0.9),
    slab_alpha = 0.8, normalize = "panels", point_size = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66) +
  scale_fill_manual(values = c("Constriction" = "#4575B4", "Dilation" = "#D73027")) +
  labs(
    x = "Posterior trigger effect (Trigger \u2013 NoTrigger, item d\u2032)",
    y = NULL,
    title = "Trigger rescue by pupil direction"
  ) +
  theme(legend.position = "none")

plot.direction_trigger
```

## Hidden trigger effect

```{r}
# Filter to the two reclassified cells of interest
df.cell_comparison <- df.v2_for_dprime |>
  filter(!trial_burnin) |>
  mutate(
    reclassified_cell = case_when(
      trigger_status_original_4cat == "lapse_notrigger" &
        trigger_status_new == "nolapse_notrigger" ~ "A_lapse_notrig_to_nolapse",
      trigger_status_original_4cat == "lapse_trigger" &
        trigger_status_new == "nolapse_trigger" ~ "B_lapse_trig_to_nolapse",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(reclassified_cell))

message("Trials per reclassified cell:")
df.cell_comparison |> count(reclassified_cell)
```

```{r}
# Item d' for hidden trigger effect
df.cell_item_dprime <- df.cell_comparison |>
  group_by(participant_id, reclassified_cell) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_target = sum(goal %in% c("BGSM", "PLUP")),
    n_foil = sum(goal == "NOV"),
    n_trials = n(),
    .groups = "drop"
  ) |>
  mutate(item_dp = fun.dprime(item_hit, item_fa, n_target, n_foil))

df.cell_item_wide <- df.cell_item_dprime |>
  select(participant_id, reclassified_cell, item_dp, n_trials) |>
  pivot_wider(names_from = reclassified_cell, values_from = c(item_dp, n_trials)) |>
  filter(!is.na(item_dp_A_lapse_notrig_to_nolapse),
         !is.na(item_dp_B_lapse_trig_to_nolapse)) |>
  mutate(delta_dp = item_dp_B_lapse_trig_to_nolapse - item_dp_A_lapse_notrig_to_nolapse)

cat("=== ITEM d': Hidden Trigger Effect ===\n")
cat(sprintf("Subjects: %d\n\n", nrow(df.cell_item_wide)))

cat(sprintf("Cell A (lapse_notrig -> nolapse): M=%.3f, SD=%.3f, trials=%.1f\n",
    mean(df.cell_item_wide$item_dp_A_lapse_notrig_to_nolapse),
    sd(df.cell_item_wide$item_dp_A_lapse_notrig_to_nolapse),
    mean(df.cell_item_wide$n_trials_A_lapse_notrig_to_nolapse)))
cat(sprintf("Cell B (lapse_trig -> nolapse):   M=%.3f, SD=%.3f, trials=%.1f\n\n",
    mean(df.cell_item_wide$item_dp_B_lapse_trig_to_nolapse),
    sd(df.cell_item_wide$item_dp_B_lapse_trig_to_nolapse),
    mean(df.cell_item_wide$n_trials_B_lapse_trig_to_nolapse)))

t_hidden_item <- t.test(
  df.cell_item_wide$item_dp_B_lapse_trig_to_nolapse,
  df.cell_item_wide$item_dp_A_lapse_notrig_to_nolapse,
  paired = TRUE, alternative = "greater"
)
print(t_hidden_item)

d_hidden_item <- effectsize::cohens_d(
  df.cell_item_wide$item_dp_B_lapse_trig_to_nolapse,
  df.cell_item_wide$item_dp_A_lapse_notrig_to_nolapse,
  paired = TRUE
)
cat(sprintf("\nCohen's d: %.3f | Mean delta: %.3f\n",
    d_hidden_item$Cohens_d, mean(df.cell_item_wide$delta_dp)))
```

## Hidden trigger effect: Bayesian models

### Item d'

```{r warning=FALSE}
df.cell_item_dprime <- df.cell_item_dprime |>
  mutate(condition = factor(reclassified_cell,
    levels = c("A_lapse_notrig_to_nolapse", "B_lapse_trig_to_nolapse"),
    labels = c("NoTrigger", "Trigger")))

fit.brm_hidden_item = brm(
  item_dp ~ 0 + condition,
  data    = df.cell_item_dprime,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_v2_hidden_trigger_item_dprime"
)

fit.brm_hidden_item |> summary()

# Trigger > NoTrigger (hidden trigger effect)
hypothesis(fit.brm_hidden_item, "conditionTrigger > conditionNoTrigger")

fit.brm_hidden_item |> emmeans(pairwise ~ condition)
```

### Source d'

```{r}
df.cell_source_dprime <- df.cell_comparison |>
  group_by(participant_id, reclassified_cell) |>
  summarise(
    src_hit = sum(resp_src_mem_acc %in% c("Size Hit", "Pleasant Hit")),
    src_fa = sum(resp_src_mem_acc %in% c("Size FA", "Pleasant FA")),
    n_src_target = sum(goal %in% c("BGSM", "PLUP")),
    n_src_foil = sum(goal == "NOV"),
    n_trials = n(),
    .groups = "drop"
  ) |>
  mutate(
    src_dp = fun.dprime(src_hit, src_fa, n_src_target, n_src_foil),
    condition = factor(reclassified_cell,
      levels = c("A_lapse_notrig_to_nolapse", "B_lapse_trig_to_nolapse"),
      labels = c("NoTrigger", "Trigger"))
  )
```

```{r warning=FALSE}
fit.brm_hidden_source = brm(
  src_dp ~ 0 + condition,
  data    = df.cell_source_dprime,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_v2_hidden_trigger_source_dprime"
)

fit.brm_hidden_source |> summary()

# Trigger > NoTrigger (hidden trigger effect for source)
hypothesis(fit.brm_hidden_source, "conditionTrigger > conditionNoTrigger")

fit.brm_hidden_source |> emmeans(pairwise ~ condition)
```

## Within-subject correlation: Original vs. hidden trigger effect

```{r}
# Compute original trigger effect per subject (LapseTrigger - Lapse item d')
df.original_trigger_effect <- df.v2_for_dprime |>
  filter(!trial_burnin) |>
  filter(trigger_status %in% c("lapse_no_trigger", "lapse_trigger")) |>
  group_by(participant_id, trigger_status) |>
  summarise(
    item_hit = sum(resp_item_mem_acc == "Item Hit"),
    item_fa = sum(resp_item_mem_acc == "Item FA"),
    n_item_target = sum(goal %in% c("BGSM", "PLUP")),
    n_item_foil = sum(goal == "NOV"),
    .groups = "drop"
  ) |>
  mutate(item_dp = fun.dprime(item_hit, item_fa, n_item_target, n_item_foil)) |>
  select(participant_id, trigger_status, item_dp) |>
  pivot_wider(names_from = trigger_status, values_from = item_dp) |>
  mutate(original_trigger_effect = lapse_trigger - lapse_no_trigger)

# Merge with hidden trigger effect
df.trigger_correlation <- df.original_trigger_effect |>
  inner_join(
    df.cell_item_wide |> select(participant_id, delta_dp),
    by = "participant_id"
  ) |>
  rename(hidden_trigger_effect = delta_dp)

message(glue("Subjects with both effects: {nrow(df.trigger_correlation)}"))

# Frequentist correlation
cor.test(df.trigger_correlation$original_trigger_effect,
         df.trigger_correlation$hidden_trigger_effect)
```

```{r warning=FALSE}
fit.brm_trigger_correlation = brm(
  hidden_trigger_effect ~ original_trigger_effect,
  data    = df.trigger_correlation,
  family  = gaussian(),
  chains  = 4,
  cores   = 4,
  seed    = 512,
  iter    = 4000,
  warmup  = 1000,
  file    = "models/brm_v2_trigger_correlation"
)

fit.brm_trigger_correlation |> summary()

# Positive slope (common mechanism)
hypothesis(fit.brm_trigger_correlation, "original_trigger_effect > 0")
```

```{r fig.width=6, fig.height=5, fig.dpi=300}
# Posterior regression lines
post_draws <- as_draws_df(fit.brm_trigger_correlation) |>
  select(b_Intercept, b_original_trigger_effect) |>
  slice_sample(n = 200)

r_val <- cor(df.trigger_correlation$original_trigger_effect,
             df.trigger_correlation$hidden_trigger_effect)

plot.trigger_correlation <- ggplot(df.trigger_correlation,
       aes(x = original_trigger_effect, y = hidden_trigger_effect)) +
  geom_abline(
    data = post_draws,
    aes(intercept = b_Intercept, slope = b_original_trigger_effect),
    alpha = 0.05, color = "steelblue"
  ) +
  geom_point(size = 2.5, alpha = 0.7) +
  geom_abline(
    intercept = fixef(fit.brm_trigger_correlation)[1, 1],
    slope = fixef(fit.brm_trigger_correlation)[2, 1],
    color = "steelblue", linewidth = 1.2
  ) +
  labs(
    x = "Original trigger effect (\u0394 item d\u2032)",
    y = "Hidden trigger effect (\u0394 item d\u2032)"
  )

plot.trigger_correlation
```

## Composite figure

```{r fig.width=14, fig.height=12, fig.dpi=300}
# Hidden trigger contrast plot (item + source in one panel)
hidden_item_draws <- as_draws_df(fit.brm_hidden_item) %>%
  transmute(
    contrast = b_conditionTrigger - b_conditionNoTrigger,
    outcome = "Item d\u2032"
  )

hidden_source_draws <- as_draws_df(fit.brm_hidden_source) %>%
  transmute(
    contrast = b_conditionTrigger - b_conditionNoTrigger,
    outcome = "Source d\u2032"
  )

hidden_draws <- bind_rows(hidden_item_draws, hidden_source_draws) %>%
  mutate(outcome = factor(outcome, levels = c("Source d\u2032", "Item d\u2032")))

plot.hidden_trigger_contrasts <- ggplot(hidden_draws, aes(x = contrast, y = outcome, fill = outcome)) +
  stat_halfeye(
    point_interval = median_hdi, .width = c(0.66, 0.9),
    slab_alpha = 0.8, normalize = "panels", point_size = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66, color = "black") +
  scale_fill_manual(values = c("Item d\u2032" = "#F67E00", "Source d\u2032" = "#FCBF4A")) +
  labs(
    x = "Posterior contrast (Trigger \u2013 Lapse)",
    y = NULL,
    title = "Hidden trigger effect"
  ) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

# Compose figure
  (plot.spline_item_memory_contrasts | plot.spline_source_memory_contrasts) /
  (plot.hidden_trigger_contrasts | plot.trigger_correlation) +
  plot_annotation(tag_levels = "a") &
  theme(plot.tag = element_text(size = 32, face = "bold", family = "Arial"))

ggsave(here("plots", "fig4_v2.svg"),
       width = 14, height = 12, dpi = 300, bg = "white")
```
