---
title: "dprime analysis"
author: "shawn schwartz"
output: html_document
date: "2026-04-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(targets)
library(knitr)
library(kableExtra)
library(showtext)
library(brms)
library(tidybayes)
library(emmeans)
library(ggdist)
library(patchwork)
library(tidyverse)

font_add_google("Nunito", "Inter", regular.wt = 300)
showtext_auto()
showtext_opts(dpi = 300)

theme_set(
  theme_classic(base_size = 24, base_family = "Inter") +
    theme(text = element_text(size = 24))
)

knitr::opts_chunk$set(comment = "", fig.show = "hold")

options(dplyr.summarise.inform = FALSE)
options(ggplot2.discrete.color = RColorBrewer::brewer.pal(9, "Set1"))

set.seed(512)

cond_levels = c(
  "no_lapse_no_trigger",
  "lapse_no_trigger",
  "lapse_trigger"
)

cond_labels = c(
  "NoLapse",
  "Lapse",
  "LapseTrigger"
)
```

# Item *d'*

```{r}
df_item = tar_read(item_dprime_trimmed, store = here("_targets")) |> 
  mutate(
    condition = factor(
      dim.trigger_condition,
      levels = cond_levels,
      labels = cond_labels
    )
  )

df_item |> head()
```

```{r fig.width=15, fig.height=15, fig.dpi=72}
fit_brm_item_dprime = brm(
  item_dp ~ 0 + dim.trigger_condition,
  data = df_item,
  family = gaussian(),
  chains = 4,
  cores = 4,
  seed = 512,
  iter = 4000,
  warmup = 1000,
  file = here("models/brm_item_dprime")
)

fit_brm_item_dprime |> summary()
fit_brm_item_dprime |> plot(nvariables = 8)
fit_brm_item_dprime |> pp_check(ndraws = 100)
fit_brm_item_dprime |> emmeans(pairwise ~ dim.trigger_condition)
```

## Lapse < NoLapse

```{r}
fit_brm_item_dprime |> 
  hypothesis(
    "dim.trigger_conditionlapse_no_trigger < dim.trigger_conditionno_lapse_no_trigger"
  )
```

## LapseTrigger > Lapse

```{r}
fit_brm_item_dprime |> 
  hypothesis(
    "dim.trigger_conditionlapse_trigger > dim.trigger_conditionlapse_no_trigger"
  )
```

## LapseTrigger < NoLapse

```{r}
fit_brm_item_dprime |> 
  hypothesis(
    "dim.trigger_conditionlapse_trigger < dim.trigger_conditionno_lapse_no_trigger"
  )
```

```{r fig.width=16, fig.height=15}
contrast_draws = fit_brm_item_dprime |> 
  spread_draws(
    b_dim.trigger_conditionno_lapse_no_trigger,
    b_dim.trigger_conditionlapse_no_trigger,
    b_dim.trigger_conditionlapse_trigger
  ) |> 
  mutate(
    `OS - SOS` = b_dim.trigger_conditionno_lapse_no_trigger - b_dim.trigger_conditionlapse_no_trigger,
    `SOS/Trigger - SOS` = b_dim.trigger_conditionlapse_trigger - b_dim.trigger_conditionlapse_no_trigger,
    `OS - SOS/Trigger` = b_dim.trigger_conditionno_lapse_no_trigger - b_dim.trigger_conditionlapse_trigger
  ) |> 
  pivot_longer(
    cols = matches("-"),
    names_to = "contrast",
    values_to = ".value"
  ) |> 
  mutate(
    contrast = factor(contrast, levels = c(
      "OS - SOS/Trigger", # bottom
      "SOS/Trigger - SOS", # middle
      "OS - SOS" # top
    ))
  )

fill_by_type = c(
  "OS - SOS" = "#D62729",
  "SOS/Trigger - SOS" = "#F67E00",
  "OS - SOS/Trigger" = "#FCBF4A"
)

plot_item_memory_contrasts = contrast_draws |> 
  ggplot(aes(x = .value, y = contrast, fill = contrast)) +
  stat_halfeye(
    point_interval = median_hdi,
    .width = c(0.66, 0.9),
    slab_alpha = 0.8,
    normalize = "panels",
    point_size = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66, color = "black") +
  geom_text(
    data = contrast_draws |> group_by(contrast) |> summarize(m = median(.value)),
    aes(x = m, y = contrast, label = contrast),
    vjust = -0.8, hjust = 0.5, size = 6, inherit.aes = FALSE, family = "Inter"
  ) +
  scale_fill_manual(values = fill_by_type) +
  scale_y_discrete(labels = NULL) +
  labs(
    x = "Posterior contrast (d\u2032)",
    y = NULL,
    title = "Item memory"
  ) +
  coord_cartesian(xlim = c(-0.75, 0.75)) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

plot_item_memory_contrasts
```

```{r}
df_item_wide = df_item |> 
  select(dim.top_level__subject_id, dim.trigger_condition, item_dp) |> 
  pivot_wider(
    id_cols = dim.top_level__subject_id,
    names_from = dim.trigger_condition,
    values_from = item_dp,
    names_glue = "{dim.trigger_condition}_{.value}"
  )

df_item_wide |> head()
```

```{r}
condition_colors = c(
  "OS" = "steelblue",
  "SOS" = "firebrick",
  "SOS/Trigger" = "darkmagenta"
)

plot_means = function(data, mem_type = c("item", "src")) {
  mem_type = match.arg(mem_type)
  dp_suffix = paste0("_", mem_type, "_dp")
  title = if (mem_type == "item") "Item memory" else "Source memory"
  
  col_nolapse = paste0("no_lapse_no_trigger", dp_suffix)
  col_lnt = paste0("lapse_no_trigger", dp_suffix)
  col_lt = paste0("lapse_trigger", dp_suffix)
  
  long = data |> 
    select(dim.top_level__subject_id, all_of(c(col_nolapse, col_lnt, col_lt))) |> 
    pivot_longer(-dim.top_level__subject_id, names_to = "condition", values_to = "dprime") |> 
    mutate(
      trigger_condition = case_when(
        condition == col_lnt ~ "SOS",
        condition == col_lt ~ "SOS/Trigger",
        condition == col_nolapse ~ "OS"
      ),
      trigger_condition = factor(trigger_condition, levels = c(
        "OS", "SOS", "SOS/Trigger"
      ))
    )
  
  # Cousinaeu-Morey within-subjects SEM
  k = n_distinct(long$trigger_condition)
  normed = long |> 
    group_by(dim.top_level__subject_id) |> 
    mutate(dprime_norm = dprime - mean(dprime, na.rm = TRUE)) |> 
    ungroup() |> 
    mutate(dprime_norm = dprime_norm + mean(long$dprime, na.rm = TRUE))
  
  ws_sem = normed |> 
    group_by(trigger_condition) |> 
    summarize(
      sem = sd(dprime_norm, na.rm = TRUE) / sqrt(sum(!is.na(dprime_norm))) * sqrt(k / (k - 1)),
      .groups = "drop"
    )
  
  means_df = long |> 
    group_by(trigger_condition) |> 
    summarize(mean_dp = mean(dprime, na.rm = TRUE), .groups = "drop") |> 
    left_join(ws_sem, by = "trigger_condition")
  
  ggplot(long, aes(x = trigger_condition, y = dprime)) +
    geom_pointrange(
      data = means_df,
      aes(x = trigger_condition, y = mean_dp,
          ymin = mean_dp - sem, ymax = mean_dp + sem,
          color = trigger_condition),
      size = 0.8, linewidth = 1
    ) +
    scale_color_manual(values = condition_colors) +
    scale_fill_manual(values = condition_colors) +
    labs(
      x = " ",
      y = "d\u2032",
      title = title
    ) +
    theme(
      axis.title.y = element_text(face = "italic"),
      legend.position = "none"
    )
}

p1 = plot_means(df_item_wide, "item")
p1
```

```{r}
make_violin_poly = function(vals, x_pos, max_width = 0.5, side = "right") {
  kde = density(vals, n = 300)
  d = kde$y / max(kde$y) * max_width
  if (side == "left") d = -d
  
  above = which(kde$x >= 0)
  below = which(kde$x <= 0)
  
  bind_rows(
    tibble(
      x = c(x_pos, x_pos + d[above], x_pos),
      y = c(kde$x[min(above)], kde$x[above], kde$x[max(above)]),
      fill = "pos"
    ),
    tibble(
      x = c(x_pos, x_pos + d[below], x_pos),
      y = c(kde$x[min(below)], kde$x[below], kde$x[max(below)]),
      fill = "neg"
    )
  )
}

plot_raincloud_diff = function(data, mem_type = c("item", "src")) {
  mem_type = match.arg(mem_type)
  dp_suffix = paste0("_", mem_type, "_dp")
  title = if (mem_type == "item") "Item memory" else "Source memory"
  
  col_nolapse = paste0("no_lapse_no_trigger", dp_suffix)
  col_lnt = paste0("lapse_no_trigger", dp_suffix)
  col_lt = paste0("lapse_trigger", dp_suffix)
  
  diff_long = data |> 
    transmute(
      dim.top_level__subject_id,
      lnt_nlnt = .data[[col_lnt]] - .data[[col_nolapse]],
      lt_lnt = .data[[col_lt]] - .data[[col_lnt]]
    ) |> 
    pivot_longer(
      -dim.top_level__subject_id, 
      names_to = "comparison", 
      values_to = "diff"
    ) |> 
    group_by(dim.top_level__subject_id) |> 
    mutate(jitter_offset = runif(1, 0.05, 0.25)) |> 
    ungroup() |> 
    mutate(
      x_pos = ifelse(comparison == "lnt_nlnt", 0, 1.5),
      sign_group = ifelse(diff >= 0, "pos", "neg"),
      x_jitter = x_pos + ifelse(comparison == "lnt_nlnt", jitter_offset, -jitter_offset)
    )
  
  violin_polys = diff_long |> 
    distinct(comparison, x_pos) |> 
    mutate(side = ifelse(comparison == "lnt_nlnt", "left", "right")) |> 
    rowwise() |> 
    mutate(
      poly = list(
        make_violin_poly(diff_long$diff[diff_long$comparison == comparison], x_pos, side = side)
      )
    ) |> 
    unnest(poly)
  
  summary_df = diff_long |> 
    group_by(comparison, x_pos) |> 
    summarize(
      mean_val = mean(diff, na.rm = TRUE),
      sem = sd(diff, na.rm = TRUE) / sqrt(sum(!is.na(diff))),
      .groups = "drop"
    )
  
  x_labels = c(
    "0" = "SOS - OS",
    "1.5" = "SOS/Trigger - SOS"
  )
  
  ggplot() +
    geom_hline(yintercept = 0, linetype = "dashed", linewidth = 1) +
    geom_polygon(
      data = violin_polys,
      aes(x = x, y = y, group = interaction(comparison, fill), fill = fill),
      alpha = 0.9,
      color = NA
    ) +
    geom_line(
      data = diff_long,
      aes(x = x_jitter, y = diff, group = dim.top_level__subject_id),
      color = "gray",
      alpha = 0.33,
      linewidth = 0.5
    ) +
    geom_point(
      data = diff_long,
      aes(x = x_jitter, y = diff, fill = sign_group),
      shape = 21,
      size = 2.5,
      alpha = 0.8,
      color = "white",
      stroke = 0.5
    ) +
    geom_segment(
      data = summary_df,
      aes(x = x_pos, xend = x_pos,
          y = mean_val - sem, yend = mean_val + sem),
      linewidth = 1.5
    ) +
    geom_segment(
      data = summary_df,
      aes(x = x_pos - 0.04, xend = x_pos + 0.04,
          y = mean_val - sem, yend = mean_val - sem),
      linewidth = 1.5
    ) +
    geom_segment(
      data = summary_df,
      aes(x = x_pos - 0.04, xend = x_pos + 0.04,
          y = mean_val + sem, yend = mean_val + sem),
      linewidth = 1.5
    ) +
    scale_fill_manual(
      values = c("pos" = "#4B999E", "neg" = "#D7753E"),
      labels = c("pos" = "Δ ≥ 0", "neg" = "Δ < 0")
    ) +
    scale_x_continuous(
      breaks = c(0, 1.5),
      labels = x_labels[c("0", "1.5")],
      limits = c(-0.8, 2.3)
    ) +
    scale_y_continuous(breaks = seq(-2, 2, 1)) +
    labs(
      x = " ",
      y = "Δd\u2032",
      title = title,
      fill = NULL
    ) +
    theme(
      axis.title.y = element_text(face = "italic"),
      legend.position = c(0.85, 0.1),
      legend.background = element_blank(),
      legend.text = element_text(size = 11)
    )
}

p2 = plot_raincloud_diff(df_item_wide, "item")
p2
```

# Source *d'*

```{r}
df_src = tar_read(src_dprime_trimmed, store = here("_targets")) |> 
  mutate(
    condition = factor(
      dim.trigger_condition,
      levels = cond_levels,
      labels = cond_labels
    )
  )

df_src |> head()
```

```{r fig.width=15, fig.height=15, fig.dpi=72}
fit_brm_src_dprime = brm(
  src_dp ~ 0 + dim.trigger_condition,
  data = df_src,
  family = gaussian(),
  chains = 4,
  cores = 4,
  seed = 512,
  iter = 4000,
  warmup = 1000,
  file = here("models/brm_src_dprime")
)

fit_brm_src_dprime |> summary()
fit_brm_src_dprime |> plot(nvariables = 8)
fit_brm_src_dprime |> pp_check(ndraws = 100)
fit_brm_src_dprime |> emmeans(pairwise ~ dim.trigger_condition)
```

## Lapse < NoLapse

```{r}
fit_brm_src_dprime |> 
  hypothesis(
    "dim.trigger_conditionlapse_no_trigger < dim.trigger_conditionno_lapse_no_trigger"
  )
```

## LapseTrigger > Lapse

```{r}
fit_brm_src_dprime |> 
  hypothesis(
    "dim.trigger_conditionlapse_trigger > dim.trigger_conditionlapse_no_trigger"
  )
```

## LapseTrigger < NoLapse

```{r}
fit_brm_src_dprime |> 
  hypothesis(
    "dim.trigger_conditionlapse_trigger < dim.trigger_conditionno_lapse_no_trigger"
  )
```

```{r fig.width=16, fig.height=15}
contrast_draws = fit_brm_src_dprime |> 
  spread_draws(
    b_dim.trigger_conditionno_lapse_no_trigger,
    b_dim.trigger_conditionlapse_no_trigger,
    b_dim.trigger_conditionlapse_trigger
  ) |> 
  mutate(
    `OS - SOS` = b_dim.trigger_conditionno_lapse_no_trigger - b_dim.trigger_conditionlapse_no_trigger,
    `SOS/Trigger - SOS` = b_dim.trigger_conditionlapse_trigger - b_dim.trigger_conditionlapse_no_trigger,
    `OS - SOS/Trigger` = b_dim.trigger_conditionno_lapse_no_trigger - b_dim.trigger_conditionlapse_trigger
  ) |> 
  pivot_longer(
    cols = matches("-"),
    names_to = "contrast",
    values_to = ".value"
  ) |> 
  mutate(
    contrast = factor(contrast, levels = c(
      "OS - SOS/Trigger", # bottom
      "SOS/Trigger - SOS", # middle
      "OS - SOS" # top
    ))
  )

fill_by_type = c(
  "OS - SOS" = "#D62729",
  "SOS/Trigger - SOS" = "#F67E00",
  "OS - SOS/Trigger" = "#FCBF4A"
)

plot_src_memory_contrasts = contrast_draws |> 
  ggplot(aes(x = .value, y = contrast, fill = contrast)) +
  stat_halfeye(
    point_interval = median_hdi,
    .width = c(0.66, 0.9),
    slab_alpha = 0.8,
    normalize = "panels",
    point_size = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66, color = "black") +
  geom_text(
    data = contrast_draws |> group_by(contrast) |> summarize(m = median(.value)),
    aes(x = m, y = contrast, label = contrast),
    vjust = -0.8, hjust = 0.5, size = 6, inherit.aes = FALSE, family = "Inter"
  ) +
  scale_fill_manual(values = fill_by_type) +
  scale_y_discrete(labels = NULL) +
  labs(
    x = "Posterior contrast (d\u2032)",
    y = NULL,
    title = "Source memory"
  ) +
  coord_cartesian(xlim = c(-0.75, 0.75)) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

plot_src_memory_contrasts
```

```{r}
df_src_wide = df_src |> 
  select(dim.top_level__subject_id, dim.trigger_condition, src_dp) |> 
  pivot_wider(
    id_cols = dim.top_level__subject_id,
    names_from = dim.trigger_condition,
    values_from = src_dp,
    names_glue = "{dim.trigger_condition}_{.value}"
  )

df_src_wide |> head()
```

```{r}
p3 = plot_means(df_src_wide, "src")
p3
```

```{r}
p4 = plot_raincloud_diff(df_src_wide, "src")
p4
```

# Figure 2

```{r fig.width=18, fig.height=12}
row1 = p1 + p2 + plot_item_memory_contrasts + plot_layout(widths = c(1, 2, 1))
row2 = p3 + p4 + plot_src_memory_contrasts + plot_layout(widths = c(1, 2, 1))

row1 / row2 +
  plot_annotation(tag_levels = "a") &
  theme(plot.tag = element_text(size = 32, face = "bold", family = "Inter"))

ggsave(here("reports/figures/fig2.pdf"), width = 20, height = 12, dpi = 300, bg = "white")
ggsave(here("reports/figures/fig2.png"), width = 20, height = 12, dpi = 300, bg = "white")
```

# Post hoc reclassification

```{r}
cond_levels = c(
  "nolapse_notrigger",
  "lapse_notrigger",
  "lapse_trigger",
  "nolapse_trigger"
)

cond_labels = c(
  "NoLapse",
  "Lapse",
  "LapseTrigger",
  "NoLapseTrigger"
)

df_trials = tar_read(detrended_data_clean_trimmed, store = here("_targets"))

df_trials |> head()
```

## Trial counts

### Original (online) classification

```{r}
df_trans_table = df_trials |> 
  filter(
    dim.top_level__detrended_status %in% c(
      "lapse_notrigger",
      "lapse_trigger",
      "nolapse_notrigger",
      "nolapse_trigger"
    )
  )

df_trans_table$dim.trigger_condition |> table()
```

### Original (online) classification vs. spline reclassified status

```{r}
table(
  df_trans_table$dim.trigger_condition,
  df_trans_table$dim.top_level__detrended_status
)
```

## Model fitting

### Item *d'*

```{r}
df_item_reclass = tar_read(item_dprime_trimmed_reclassified, store = here("_targets")) |> 
  mutate(
    condition = factor(
      dim.top_level__detrended_status,
      levels = cond_levels,
      labels = cond_labels
    )
  )

df_item_reclass |> head()

table(
  df_item_reclass$condition,
  df_item_reclass$dim.top_level__detrended_status
)
```

```{r fig.width=15, fig.height=15, fig.dpi=72}
fit_brm_item_dprime_reclassified = brm(
  item_dp ~ 0 + condition,
  data = df_item_reclass,
  family = gaussian(),
  chains = 4,
  cores = 4,
  seed = 512,
  iter = 4000,
  warmup = 1000,
  file = here("models/brm_item_dprime_reclassified")
)

fit_brm_item_dprime_reclassified |> summary()
fit_brm_item_dprime_reclassified |> plot(nvariables = 8)
fit_brm_item_dprime_reclassified |> pp_check(ndraws = 100)
fit_brm_item_dprime_reclassified |> emmeans(pairwise ~ condition)
```

#### Lapse < NoLapse

```{r}
fit_brm_item_dprime_reclassified |> 
  hypothesis("conditionLapse < conditionNoLapse")
```

#### LapseTrigger > Lapse

```{r}
fit_brm_item_dprime_reclassified |> 
  hypothesis("conditionLapseTrigger > conditionLapse")
```

#### LapseTrigger < NoLapse

```{r}
fit_brm_item_dprime_reclassified |> 
  hypothesis("conditionLapseTrigger < conditionNoLapse")
```

#### Posterior contrasts plot

```{r fig.width=16, fig.height=15}
contrast_draws = fit_brm_item_dprime_reclassified |> 
  spread_draws(
    b_conditionNoLapse,
    b_conditionLapse,
    b_conditionLapseTrigger
  ) |> 
  mutate(
    `OS - SOS` = b_conditionNoLapse - b_conditionLapse,
    `SOS/Trigger - SOS` = b_conditionLapseTrigger - b_conditionLapse,
    `OS - SOS/Trigger` = b_conditionNoLapse - b_conditionLapseTrigger
  ) |> 
  pivot_longer(
    cols = matches("-"),
    names_to = "contrast",
    values_to = ".value"
  ) |> 
  mutate(
    contrast = factor(contrast, levels = c(
      "OS - SOS/Trigger", # bottom
      "SOS/Trigger - SOS", # middle
      "OS - SOS" # top
    ))
  )

fill_by_type = c(
  "OS - SOS" = "#D62729",
  "SOS/Trigger - SOS" = "#F67E00",
  "OS - SOS/Trigger" = "#FCBF4A"
)

plot_item_memory_contrasts_reclassified = contrast_draws |> 
  ggplot(aes(x = .value, y = contrast, fill = contrast)) +
  stat_halfeye(
    point_interval = median_hdi,
    .width = c(0.66, 0.9),
    slab_alpha = 0.8,
    normalize = "panels",
    point_size = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66, color = "black") +
  geom_text(
    data = contrast_draws |> group_by(contrast) |> summarize(m = median(.value)),
    aes(x = m, y = contrast, label = contrast),
    vjust = -0.8, hjust = 0.5, size = 6, inherit.aes = FALSE, family = "Inter"
  ) +
  scale_fill_manual(values = fill_by_type) +
  scale_y_discrete(labels = NULL) +
  labs(
    x = "Posterior contrast (d\u2032)",
    y = NULL,
    title = "Item memory (reclassified)"
  ) +
  coord_cartesian(xlim = c(-0.75, 0.75)) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

plot_item_memory_contrasts_reclassified
```

### Source *d'*

```{r}
df_src_reclass = tar_read(src_dprime_trimmed_reclassified, store = here("_targets")) |> 
  mutate(
    condition = factor(
      dim.top_level__detrended_status,
      levels = cond_levels,
      labels = cond_labels
    )
  )

df_src_reclass |> head()

table(
  df_src_reclass$condition,
  df_src_reclass$dim.top_level__detrended_status
)
```

```{r fig.width=15, fig.height=15, fig.dpi=72}
fit_brm_src_dprime_reclassified = brm(
  src_dp ~ 0 + condition,
  data = df_src_reclass,
  family = gaussian(),
  chains = 4,
  cores = 4,
  seed = 512,
  iter = 4000,
  warmup = 1000,
  file = here("models/brm_src_dprime_reclassified")
)

fit_brm_src_dprime_reclassified |> summary()
fit_brm_src_dprime_reclassified |> plot(nvariables = 8)
fit_brm_src_dprime_reclassified |> pp_check(ndraws = 100)
fit_brm_src_dprime_reclassified |> emmeans(pairwise ~ condition)
```

#### Lapse < NoLapse

```{r}
fit_brm_src_dprime_reclassified |> 
  hypothesis("conditionLapse < conditionNoLapse")
```

#### LapseTrigger > Lapse

```{r}
fit_brm_src_dprime_reclassified |> 
  hypothesis("conditionLapseTrigger > conditionLapse")
```

#### LapseTrigger < NoLapse

```{r}
fit_brm_src_dprime_reclassified |> 
  hypothesis("conditionLapseTrigger < conditionNoLapse")
```

#### Posterior contrasts plot

```{r fig.width=16, fig.height=15}
contrast_draws = fit_brm_src_dprime_reclassified |> 
  spread_draws(
    b_conditionNoLapse,
    b_conditionLapse,
    b_conditionLapseTrigger
  ) |> 
  mutate(
    `OS - SOS` = b_conditionNoLapse - b_conditionLapse,
    `SOS/Trigger - SOS` = b_conditionLapseTrigger - b_conditionLapse,
    `OS - SOS/Trigger` = b_conditionNoLapse - b_conditionLapseTrigger
  ) |> 
  pivot_longer(
    cols = matches("-"),
    names_to = "contrast",
    values_to = ".value"
  ) |> 
  mutate(
    contrast = factor(contrast, levels = c(
      "OS - SOS/Trigger", # bottom
      "SOS/Trigger - SOS", # middle
      "OS - SOS" # top
    ))
  )

fill_by_type = c(
  "OS - SOS" = "#D62729",
  "SOS/Trigger - SOS" = "#F67E00",
  "OS - SOS/Trigger" = "#FCBF4A"
)

plot_src_memory_contrasts_reclassified = contrast_draws |> 
  ggplot(aes(x = .value, y = contrast, fill = contrast)) +
  stat_halfeye(
    point_interval = median_hdi,
    .width = c(0.66, 0.9),
    slab_alpha = 0.8,
    normalize = "panels",
    point_size = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66, color = "black") +
  geom_text(
    data = contrast_draws |> group_by(contrast) |> summarize(m = median(.value)),
    aes(x = m, y = contrast, label = contrast),
    vjust = -0.8, hjust = 0.5, size = 6, inherit.aes = FALSE, family = "Inter"
  ) +
  scale_fill_manual(values = fill_by_type) +
  scale_y_discrete(labels = NULL) +
  labs(
    x = "Posterior contrast (d\u2032)",
    y = NULL,
    title = "Source memory (reclassified)"
  ) +
  coord_cartesian(xlim = c(-0.75, 0.75)) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

plot_src_memory_contrasts_reclassified
```

## Effect contrasts

Here, we create a 6-cell variable crossing online classification and reclassified status
so that we can test contrastss on specific cells of the transition table.

```{r}
source(here("R/derive_dprime.R"))

df_item_cells = df_trials |> 
  filter(
    !dim.top_level__burn_in_trial,
    !dim.ret_response_item_acc %in% c("blink", "NO_RESPONSE"),
    !dim.top_level__detrended_status %in% "NA_notrigger"
  ) |> 
  mutate(
    cell = paste(dim.trigger_condition, dim.top_level__detrended_status, sep = "__")
  ) |> 
  group_by(dim.top_level__subject_id, cell) |> 
  summarise(
    item_hit       = sum(dim.ret_response_item_acc == "Item Hit"),
    item_miss      = sum(dim.ret_response_item_acc == "Item Miss"),
    item_fa        = sum(dim.ret_response_item_acc == "Item FA"),
    item_cr        = sum(dim.ret_response_item_acc == "Item CR"),
    n_item_target  = sum(dim.top_level__enc_goal %in% c("BGSM", "PLUP")),
    n_item_foil    = sum(dim.top_level__enc_goal == "NOV"),
    n_item_total   = item_hit + item_miss + item_fa + item_cr,
    .groups        = "drop"
  ) %>%
  mutate(
    item_dp  = calc_dprime(item_hit, item_fa, n_item_target, n_item_foil),
    cell = factor(cell, levels = c(
      "lapse_no_trigger__lapse_notrigger",    # 2011
      "no_lapse_no_trigger__lapse_notrigger", # 511
      "lapse_no_trigger__nolapse_notrigger",  # 2466
      "no_lapse_no_trigger__nolapse_notrigger", # 3626
      "lapse_trigger__lapse_trigger",         # 2016
      "lapse_trigger__nolapse_trigger"        # 2451
    ))
  )

table(df_item_cells$cell)
```

```{r fig.width=15, fig.height=15, fig.dpi=72}
fit_brm_item_dprime_cells = brm(
  item_dp ~ 0 + cell,
  data = df_item_cells,
  family = gaussian(),
  chains = 4,
  cores = 4,
  seed = 512,
  iter = 4000,
  warmup = 1000,
  file = here("models/brm_item_dprime_cells")
)

fit_brm_item_dprime_cells |> summary()
fit_brm_item_dprime_cells |> plot(nvariables = 8)
fit_brm_item_dprime_cells |> pp_check(ndraws = 100)
fit_brm_item_dprime_cells |> emmeans(pairwise ~ cell)
```

### Cell-level *d'* ranking

```{r fig.width=10, fig.height=10}
cell_labels = c(
  "no_lapse_no_trigger__lapse_notrigger" = "Online OS -> SOS", #"511",
  "lapse_no_trigger__lapse_notrigger" = "Drift-resistant SOS", #"2011",
  "lapse_no_trigger__nolapse_notrigger" = "Drift-driven SOS", #"2466",
  "lapse_trigger__lapse_trigger" = "SOS-triggered", #"2016",
  "lapse_trigger__nolapse_trigger" = "SOS-triggered -> OS", #"2451",
  "no_lapse_no_trigger__nolapse_notrigger" = "Steadfast OS" #"3626"
)

# order by posterior mean d'
cell_order = fit_brm_item_dprime_cells |> 
  spread_draws(
    b_celllapse_no_trigger__lapse_notrigger,
    b_cellno_lapse_no_trigger__lapse_notrigger,
    b_celllapse_no_trigger__nolapse_notrigger,
    b_celllapse_trigger__lapse_trigger,
    b_celllapse_trigger__nolapse_trigger,
    b_cellno_lapse_no_trigger__nolapse_notrigger
  ) |> 
  pivot_longer(starts_with("b_cell"), names_to = "cell", values_to = "dp") |> 
  mutate(cell = str_remove(cell, "^b_cell")) |> 
  group_by(cell) |> 
  summarise(mean_dp = mean(dp)) |> 
  arrange(mean_dp) |> 
  pull(cell)

cell_draws = fit_brm_item_dprime_cells |> 
  spread_draws(
    b_celllapse_no_trigger__lapse_notrigger,
    b_cellno_lapse_no_trigger__lapse_notrigger,
    b_celllapse_no_trigger__nolapse_notrigger,
    b_celllapse_trigger__lapse_trigger,
    b_celllapse_trigger__nolapse_trigger,
    b_cellno_lapse_no_trigger__nolapse_notrigger
  ) |> 
  pivot_longer(starts_with("b_cell"), names_to = "cell", values_to = "dp") |> 
  mutate(
    cell = str_remove(cell, "^b_cell"),
    cell = factor(cell, levels = cell_order),
    label = cell_labels[as.character(cell)]
  ) 

cell_colors = c(
  "no_lapse_no_trigger__lapse_notrigger" = "#C84D4C", # 511
  "lapse_no_trigger__lapse_notrigger" = "#D37538", # 2011
  "lapse_no_trigger__nolapse_notrigger" = "#CE9A28", # 2466
  "lapse_trigger__lapse_trigger" = "#9163C1", # 2016
  "lapse_trigger__nolapse_trigger" = "#3F78C1", # 2451
  "no_lapse_no_trigger__nolapse_notrigger" = "#356920" # 3626
)

plot_dprime_bars = ggplot(cell_draws, aes(x = reorder(label, dp, FUN = median), y = dp, fill = cell)) + 
  stat_summary(fun = mean, geom = "bar", width = 0.4, alpha = 1.0) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 0.7) +
  scale_fill_manual(values = cell_colors, guide = "none") +
  scale_y_continuous(expand = c(0,0)) + 
  labs(
    x = NULL,
    y = "Item d\u2032"
    # title = "Item d\u2032 by online × reclassified cell"
  ) +
  theme(
    # text = element_text(size = 8),
    # axis.text.x = element_text(size = 8, lineheight = 0.9, angle = 90, vjust = 0.5),
    # axis.text.x = element_text(lineheight = 0.9, angle = 90, vjust = 0.5),
    legend.position = "none"
  ) +
  coord_flip()

plot_dprime_bars
```

### Pooled lapse effect: lapse_notrigger < nolapse_notrigger

```{r}
# (2011 + 511) < (2466 + 3626)
fit_brm_item_dprime_cells |> 
  hypothesis(
    "(celllapse_no_trigger__lapse_notrigger + cellno_lapse_no_trigger__lapse_notrigger) / 2 < (celllapse_no_trigger__nolapse_notrigger + cellno_lapse_no_trigger__nolapse_notrigger) / 2"
  )
```

### Within online lapse_no_trigger: reclassified lapse < reclassified nolapse

```{r}
# 2011 < 2466
fit_brm_item_dprime_cells |> 
  hypothesis(
    "celllapse_no_trigger__lapse_notrigger < celllapse_no_trigger__nolapse_notrigger"
  )
```

### Trigger effect

```{r}
# pooled lapse_notrigger < lapse_trigger
# (2011 + 511) < 2016
fit_brm_item_dprime_cells |> 
  hypothesis(
    "(celllapse_no_trigger__lapse_notrigger + cellno_lapse_no_trigger__lapse_notrigger) / 2 < celllapse_trigger__lapse_trigger"
  )
```

```{r}
# online lapse untriggered (reclassified lapse) < online lapse triggered (reclassified lapse)
# 2011 < 2016
fit_brm_item_dprime_cells |> 
  hypothesis(
    "celllapse_no_trigger__lapse_notrigger < celllapse_trigger__lapse_trigger"
  )
```

```{r}
# pooled nolapse_notrigger < nolapse_trigger
# (2466 + 3626) < 2451
fit_brm_item_dprime_cells |> 
  hypothesis(
    "(celllapse_no_trigger__nolapse_notrigger + cellno_lapse_no_trigger__nolapse_notrigger) / 2 < celllapse_trigger__nolapse_trigger"
  )
```

```{r}
# online lapse untriggered (reclassified nolapse) < online lapse triggered (reclassified nolapse)
# 2466 < 2451
fit_brm_item_dprime_cells |> 
  hypothesis(
    "celllapse_no_trigger__nolapse_notrigger < celllapse_trigger__nolapse_trigger"
  )
```

```{r}
# online no-lapse untriggered (reclassified nolapse) > online lapse triggered (reclassified nolapse)
# 3626 > 2451
fit_brm_item_dprime_cells |> 
  hypothesis(
    "cellno_lapse_no_trigger__nolapse_notrigger > celllapse_trigger__nolapse_trigger"
  )
```

```{r}
# online nolapse untriggered (reclassified nolapse) > online lapse triggered (reclassified lapse)
# 3626 > 2016
fit_brm_item_dprime_cells |> 
  hypothesis(
    "cellno_lapse_no_trigger__nolapse_notrigger > celllapse_trigger__lapse_trigger"
  )
```

```{r}
# online lapse triggered (reclsasified lapse) < online lapse triggered (reclassified nolapse)
# 2016 < 2451
fit_brm_item_dprime_cells |> 
  hypothesis(
    "celllapse_trigger__lapse_trigger < celllapse_trigger__nolapse_trigger"
  )
```

### Interaction

> Is the trigger effect larger for reclassified nolapse than reclassified lapse?

```{r}
# interaction: trigger effect for nolapse > trigger effeeect for lapse
# (2451 - 2466) > (2016 - 2011)
fit_brm_item_dprime_cells |> 
  hypothesis(
    "(celllapse_trigger__nolapse_trigger - celllapse_no_trigger__nolapse_notrigger) > (celllapse_trigger__lapse_trigger - celllapse_no_trigger__lapse_notrigger)"
  )
```


```{r fig.width=15, fig.height=15, fig.dpi=72}
# assign numeric ranks to the 4 online-SOS cells ordered by inferred depth
df_item_slope = df_item_cells |> 
  filter(cell %in% c(
    "lapse_no_trigger__lapse_notrigger", # 2011 drift-resistant, untriggered
    "lapse_no_trigger__nolapse_notrigger", # 2466 drift-driven, untriggered
    "lapse_trigger__lapse_trigger", # 2016 drift-resistant, triggered
    "lapse_trigger__nolapse_trigger" # 2451 drift-driven, triggered
  )) |> 
  mutate(depth_rank = case_when(
    cell == "lapse_no_trigger__lapse_notrigger" ~ 1,
    cell == "lapse_no_trigger__nolapse_notrigger" ~ 2,
    cell == "lapse_trigger__lapse_trigger" ~ 3,
    cell == "lapse_trigger__nolapse_trigger" ~ 4
  )) |> 
  mutate(subid = dim.top_level__subject_id)

fit_brm_item_dprime_slope = brm(
  item_dp ~ depth_rank + (1 | subid),
  data = df_item_slope,
  family = gaussian(),
  chains = 4,
  cores = 4,
  seed = 512,
  iter = 4000,
  warmup = 1000,
  file = here("models/brm_item_dprime_slope")
)


fit_brm_item_dprime_slope |> summary()
fit_brm_item_dprime_slope |> plot(nvariables = 8)
fit_brm_item_dprime_slope |> pp_check(ndraws = 100)
fit_brm_item_dprime_slope |> hypothesis("depth_rank > 0")
fit_brm_item_dprime_slope |> posterior_summary(pars = "b_depth_rank")

plot_4cell_slope = fit_brm_item_dprime_slope |> 
  spread_draws(b_depth_rank) |> 
  ggplot(aes(x = b_depth_rank)) +
  stat_halfeye(
    fill = "steelblue",
    alpha = 0.7,
    point_interval = median_hdi,
    .width = c(0.66, 0.9),
    slab_alpha = 0.8,
    normalize = "panels",
    point_size = 4
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.66, color = "black") +
  labs(
    x = "Posterior slope (Δd\u2032 per step)",
    y = NULL
    # title = "Linear trend across 4 SOS cells"
  )

plot_4cell_slope
```

## Spline drift

```{r fig.width=30, fig.height=15}
run_labels = c(
  "01" = "Run 1",
  "02" = "Run 2",
  "03" = "Run 3"
)

spline_fits = tar_read(spline_fits, store = here("_targets")) |> 
  group_by(subject, run) |> 
  mutate(
    spline_pred_mc = spline_pred - mean(spline_pred, na.rm = TRUE)
  ) |> 
  ungroup() |> 
  mutate(run_label = run_labels[run])

avg_spline = tar_read(avg_spline, store = here("_targets")) |> 
  mutate(run_label = run_labels[run])


avg_spline_mc = spline_fits |> 
  group_by(run, run_label, run_sample) |> 
  summarise(
    mean_time = mean(time, na.rm = TRUE),
    mean_spline = mean(spline_pred_mc, na.rm = TRUE),
    se_spline = sd(spline_pred_mc, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

plot_spline_spaghetti = ggplot() +
  geom_line(
    data = spline_fits,
    aes(x = time, y = spline_pred_mc, group = subject),
    alpha = 0.25, linewidth = 0.3, color = "grey50"
  ) +
  geom_ribbon(
    data = avg_spline_mc,
    aes(x = mean_time, ymin = mean_spline - se_spline, ymax = mean_spline + se_spline),
    alpha = 0.3, fill = "firebrick"
  ) +
  geom_line(
    data = avg_spline_mc,
    aes(x = mean_time, y = mean_spline),
    color = "firebrick", linewidth = 1.2
  ) +
  facet_wrap(~run_label, scales = "free_x") +
  labs(
    x = "Time (s)",
    y = "Spline-predicted pupil (a.u.)"
  ) +
  coord_cartesian(
    xlim = c(0, 800)
  )

plot_spline_spaghetti
```

# Figure 3

```{r fig.width=20, fig.height=15}
row1 = plot_spline_spaghetti + plot_item_memory_contrasts_reclassified + plot_src_memory_contrasts_reclassified + plot_layout(widths = c(2, 1, 1))
row2 = plot_spacer() + plot_dprime_bars + plot_4cell_slope + plot_spacer() + plot_layout(widths = c(1, 2, 2, 2))

row1 / row2 +
  plot_annotation(tag_levels = "a") &
  theme(plot.tag = element_text(size = 32, face = "bold", family = "Inter"))

ggsave(here("reports/figures/fig3.pdf"), width = 25, height = 13, dpi = 300, bg = "white")
ggsave(here("reports/figures/fig3.png"), width = 25, height = 13, dpi = 300, bg = "white")
```

# Asymmetry test

## Deep lapses

```{r}
# build d' per cell x constriction/dilation for deep lapse trials (2011 & 2016)

df_item_asym = df_trials |> 
  filter(
    !dim.top_level__burn_in_trial,
    !dim.ret_response_item_acc %in% c("blink", "NO_RESPONSE"),
    !dim.top_level__detrended_status %in% "NA_notrigger",
    (dim.trigger_condition == "lapse_no_trigger" & dim.top_level__detrended_status == "lapse_notrigger") |
      (dim.trigger_condition == "lapse_trigger" & dim.top_level__detrended_status == "lapse_trigger"),
    dim.realtime_window_status %in% c("constriction", "dilation")
  ) |> 
  mutate(
    cell = paste(dim.trigger_condition, dim.realtime_window_status, sep = "__")
  ) |> 
  group_by(dim.top_level__subject_id, cell) |> 
  summarise(
    item_hit       = sum(dim.ret_response_item_acc == "Item Hit"),
    item_miss      = sum(dim.ret_response_item_acc == "Item Miss"),
    item_fa        = sum(dim.ret_response_item_acc == "Item FA"),
    item_cr        = sum(dim.ret_response_item_acc == "Item CR"),
    n_item_target  = sum(dim.top_level__enc_goal %in% c("BGSM", "PLUP")),
    n_item_foil    = sum(dim.top_level__enc_goal == "NOV"),
    n_item_total   = item_hit + item_miss + item_fa + item_cr,
    .groups        = "drop"
  ) %>%
  mutate(
    item_dp  = calc_dprime(item_hit, item_fa, n_item_target, n_item_foil),
    cell = factor(cell, levels = c(
      "lapse_no_trigger__constriction",    # 2011
      "lapse_no_trigger__dilation",    # 2011
      "lapse_trigger__constriction",         # 2016
      "lapse_trigger__dilation"         # 2016
    ))
  )

table(df_item_asym$cell)
```

```{r fig.width=15, fig.height=15, fig.dpi=72}
fit_brm_item_dprime_asym = brm(
  item_dp ~ 0 + cell,
  data = df_item_asym,
  family = gaussian(),
  chains = 4,
  cores = 4,
  seed = 512,
  iter = 4000,
  warmup = 1000,
  file = here("models/brm_item_dprime_asym")
)

fit_brm_item_dprime_asym |> summary()
fit_brm_item_dprime_asym |> plot(nvariables = 8)
fit_brm_item_dprime_asym |> pp_check(ndraws = 100)
fit_brm_item_dprime_asym |> emmeans(pairwise ~ cell)
```

```{r}
# trigger boost for constrictions (2011 constrictoin > 2016 constriction)
fit_brm_item_dprime_asym |> 
  hypothesis(
    "celllapse_trigger__constriction > celllapse_no_trigger__constriction"
  )
```

```{r}
# trigger boost for dilations (2011 dilation -> 2016 dilation)
fit_brm_item_dprime_asym |> 
  hypothesis(
    "celllapse_trigger__dilation > celllapse_no_trigger__dilation"
  )
```

```{r}
# assymetry interaction: is the trigger boost larger for constrictions than dilations?
# (trigger_con - notrigger_con) > (trigger_dil - notrigger_dil)
fit_brm_item_dprime_asym |> 
  hypothesis(
    "(celllapse_trigger__constriction - celllapse_no_trigger__constriction) > (celllapse_trigger__dilation - celllapse_no_trigger__dilation)"
  )
```

## All online SOS-trials

```{r}
# all online-lapse trials (triggered and untriggered) x constriction/dilation

df_item_asym_all = df_trials |> 
  filter(
    !dim.top_level__burn_in_trial,
    !dim.ret_response_item_acc %in% c("blink", "NO_RESPONSE"),
    !dim.top_level__detrended_status %in% "NA_notrigger",
    dim.trigger_condition %in% c("lapse_no_trigger", "lapse_trigger"),
    dim.realtime_window_status %in% c("constriction", "dilation")
  ) |> 
  mutate(
    triggered = if_else(dim.trigger_condition == "lapse_trigger", "triggered", "untriggered"),
    cell = paste(triggered, dim.realtime_window_status, sep = "__")
  ) |> 
  group_by(dim.top_level__subject_id, cell) |> 
  summarise(
    item_hit       = sum(dim.ret_response_item_acc == "Item Hit"),
    item_miss      = sum(dim.ret_response_item_acc == "Item Miss"),
    item_fa        = sum(dim.ret_response_item_acc == "Item FA"),
    item_cr        = sum(dim.ret_response_item_acc == "Item CR"),
    n_item_target  = sum(dim.top_level__enc_goal %in% c("BGSM", "PLUP")),
    n_item_foil    = sum(dim.top_level__enc_goal == "NOV"),
    n_item_total   = item_hit + item_miss + item_fa + item_cr,
    .groups        = "drop"
  ) %>%
  mutate(
    item_dp  = calc_dprime(item_hit, item_fa, n_item_target, n_item_foil),
    cell = factor(cell, levels = c(
      "untriggered__constriction",    
      "untriggered__dilation",    
      "triggered__constriction",         
      "triggered__dilation"         
    ))
  )

table(df_item_asym_all$cell)
```

```{r fig.width=15, fig.height=15, fig.dpi=72}
fit_brm_item_dprime_asym_all = brm(
  item_dp ~ 0 + cell,
  data = df_item_asym_all,
  family = gaussian(),
  chains = 4,
  cores = 4,
  seed = 512,
  iter = 4000,
  warmup = 1000,
  file = here("models/brm_item_dprime_asym_all_lapse")
)

fit_brm_item_dprime_asym_all |> summary()
fit_brm_item_dprime_asym_all |> plot(nvariables = 8)
fit_brm_item_dprime_asym_all |> pp_check(ndraws = 100)
fit_brm_item_dprime_asym_all |> emmeans(pairwise ~ cell)
```

```{r}
# trigger boost for constrictions (all lapse trials)
fit_brm_item_dprime_asym_all |> 
  hypothesis(
    "celltriggered__constriction > celluntriggered__constriction"
  )
```

```{r}
# trigger boost for dilations (all lapse trials)
fit_brm_item_dprime_asym_all |> 
  hypothesis(
    "celltriggered__dilation > celluntriggered__dilation"
  )
```

```{r}
# asymmetry interaction (all lapse): trigger boost larger for constrictions than dilations?
fit_brm_item_dprime_asym_all |> 
  hypothesis(
    "(celltriggered__constriction - celluntriggered__constriction) > (celltriggered__dilation - celluntriggered__dilation)"
  )
```

## Driven-driven SOS only

```{r}
df_item_asym_moderate = df_trials |> 
  filter(
    !dim.top_level__burn_in_trial,
    !dim.ret_response_item_acc %in% c("blink", "NO_RESPONSE"),
    !dim.top_level__detrended_status %in% "NA_notrigger",
    (dim.trigger_condition == "lapse_no_trigger" & dim.top_level__detrended_status == "nolapse_notrigger") |
      (dim.trigger_condition == "lapse_trigger" & dim.top_level__detrended_status == "nolapse_trigger"),
    dim.realtime_window_status %in% c("constriction", "dilation")
  ) |> 
  mutate(
    triggered = if_else(dim.trigger_condition == "lapse_trigger", "triggered", "untriggered"),
    cell = paste(triggered, dim.realtime_window_status, sep = "__")
  ) |> 
  group_by(dim.top_level__subject_id, cell) |> 
  summarise(
    item_hit       = sum(dim.ret_response_item_acc == "Item Hit"),
    item_miss      = sum(dim.ret_response_item_acc == "Item Miss"),
    item_fa        = sum(dim.ret_response_item_acc == "Item FA"),
    item_cr        = sum(dim.ret_response_item_acc == "Item CR"),
    n_item_target  = sum(dim.top_level__enc_goal %in% c("BGSM", "PLUP")),
    n_item_foil    = sum(dim.top_level__enc_goal == "NOV"),
    n_item_total   = item_hit + item_miss + item_fa + item_cr,
    .groups        = "drop"
  ) %>%
  mutate(
    item_dp  = calc_dprime(item_hit, item_fa, n_item_target, n_item_foil),
    cell = factor(cell, levels = c(
      "untriggered__constriction",
      "untriggered__dilation",    
      "triggered__constriction",   
      "triggered__dilation"        
    ))
  )

table(df_item_asym_moderate$cell)
```

```{r fig.width=15, fig.height=15, fig.dpi=72}
fit_brm_item_dprime_asym_moderate = brm(
  item_dp ~ 0 + cell,
  data = df_item_asym_moderate,
  family = gaussian(),
  chains = 4,
  cores = 4,
  seed = 512,
  iter = 4000,
  warmup = 1000,
  file = here("models/brm_item_dprime_asym_moderate")
)

fit_brm_item_dprime_asym_moderate |> summary()
fit_brm_item_dprime_asym_moderate |> plot(nvariables = 8)
fit_brm_item_dprime_asym_moderate |> pp_check(ndraws = 100)
fit_brm_item_dprime_asym_moderate |> emmeans(pairwise ~ cell)
```

```{r}
# trigger boost for constrictions (drift-driven SOS)
fit_brm_item_dprime_asym_moderate |> 
  hypothesis(
    "celltriggered__constriction > celluntriggered__constriction"
  )
```

```{r}
# trigger boost for dilations (drift-driven SOS)
fit_brm_item_dprime_asym_moderate |> 
  hypothesis(
    "celltriggered__dilation > celluntriggered__dilation"
  )
```

```{r}
# asymmetry interaction (drift-driven SOS): trigger boost larger for constrictions than dilations?
fit_brm_item_dprime_asym_moderate |> 
  hypothesis(
    "(celltriggered__constriction - celluntriggered__constriction) > (celltriggered__dilation - celluntriggered__dilation)"
  )
```

